# Arthropod metabarcoding of surface eDNA samples
40 eDNA samples were taken from tree bark ('roller aggregation') and understory foliage ('spray aggregation') in Colliers Mills Wildlife Management Area, New Jersey, USA and processed using COI and 16S metabarcoding. This code allows reproduction of analyses, tables, and figures for the paper Allen et al. (2023). If you run the first chunk of code, you should be able to then skip ahead and run any others out of order. The sections are organized by which figure or table they create. Note that the ecotag output and MOTU tables (the first files to be read in) were created using pinebugs_post_process_final.Rmd.

Allen, M. C., Lockwood, J. L., Kwait, R., Vastano, A., Peterson, D. L. Tkacenko, L. A., … Jaffe, B. D. (2023). Using surface environmental DNA to assess arthropod biodiversity within a forested ecosystem. Environmental DNA.

# Load libraries. Read, clean and format data.
```{r}
library(dplyr) # dplyr_1.1.2 
library(readxl) # readxl_1.4.3
library(ggplot2) # ggplot2_3.4.2
library(stringr) # stringr_1.5.0
library(BiodiversityR) # BiodiversityR_2.15-3
library(wesanderson) # wesanderson_0.3.6
library(patchwork) # patchwork_1.1.2 
library(tidyr) # tidyr_1.3.0
library(iNEXT) # iNEXT_3.0.0
library(forcats) # forcats_1.0.0
select <- dplyr::select
'%notin%' <- Negate('%in%')

# read in Zeale (COI) & Coleop (16S) MOTU data with taxonomy
  # excludes MOTUs that were filtered out during negative subtraction etc.
  # these csv files were created using the pinebugs_post_process_final.Rmd script
zs3 <- read.csv("data/pinebugs.samples.onlyarth.negsub.Z_allMOTUs.csv") %>%
  filter(sampsum != 0) # removes 3 with no reads in the samples

cs3 <- read.csv("data/pinebugs.samples.onlyarth.negsub.C_allMOTUs.csv") %>%
  filter(sampsum != 0) # removes 12 with no reads in the samples

# read in family trait info (life mode)
    # terrestrial/aquatic info from Horren et al. database & other sources
    # ITT - insect trait tool (Horren et al.), while IB = Insect Biodiversity book (see Supporting Info)
lifemode <- read.csv("data/lifemode.csv")

# read in family trait info (diet/feeding)
    # diet/feeding info from Horren et al. database & other sources (see Supporting Info)
diet <- read.csv("data/diet.csv")

# read in family trait info (body size)
    # size info from Borror & Delong textbook & other sources (see Supporting Info)
size <- read.csv("data/size.csv")

# create full merged data set of arthropod MOTUs (both primer sets)
art_full <- cs3 %>%
  mutate(primer = "coleop") %>%
  bind_rows(mutate(zs3, primer = "zeale")) %>%
  select(primer, id, phylum, class, order, family, 
         genus, species, contains("CS")[c(2:10,1)], contains("DS")[c(2:7,1)],
         contains("CR")[c(2:10,1)], contains("DR")[c(2:10,1)], sampsum) %>%
  rename(sample.sum = sampsum) %>%
  mutate(sample.3CS = case_when(primer == "zeale" ~ 0,
                                  TRUE ~ sample.3CS))

# rename columns
new_col_names <- art_full %>% 
  select(starts_with("sample")) %>%
  colnames() %>%
  substr(., 1, 11) %>%
  gsub(., pattern = ":", replacement = "") %>%
  gsub(., pattern = "B", replacement = "") %>%
  gsub(., pattern = "Z", replacement = "")

colnames(art_full) # check which columns are being replaced

colnames(art_full)[9:46] <- c(new_col_names) # assign new column names

colnames(art_full) # check that it worked

# look at read depth
sort(colSums(select(filter(art_full, primer == "coleop"), starts_with("sample")), na.rm = T)) 
sort(colSums(select(filter(art_full, primer == "zeale"), starts_with("sample")), na.rm = T)) 

# calcultate mean and SE read depth by primer set
mean(colSums(select(filter(art_full, primer == "coleop"), starts_with("sample"), -sample.sum), na.rm = T))
# 60546.84

mean(sort(colSums(select(filter(art_full, primer == "zeale"), starts_with("sample"), -sample.sum), na.rm = T))[2:37])
# 114364.9

csd <- sd(colSums(select(filter(art_full, primer == "coleop"), starts_with("sample"), -sample.sum), na.rm = T))

zsd <- sd(sort(colSums(select(filter(art_full, primer == "zeale"), starts_with("sample"), -sample.sum), na.rm = T))[2:37])

cn <- sum(colSums(select(filter(art_full, primer == "coleop"), starts_with("sample"), -sample.sum), na.rm = T)>0)

zn <- sum(colSums(select(filter(art_full, primer == "zeale"), starts_with("sample"), -sample.sum), na.rm = T)>0)

(cse <- csd/sqrt(cn)) # 5392.293 # SE for 16S

(zse <- zsd/sqrt(zn)) # 6451.267 # SE for COI

# make family-lineage lookup table
lineage_lookup <- art_full %>%
  dplyr::select(phylum, class, order, family, genus, species) %>%
  distinct()

# format MOTU-level database for use in analyses
art_motu_reads <- art_full %>%
  select(primer, id, starts_with("sample")) %>%
  pivot_wider(names_from = primer, values_from = starts_with("sample")) %>%
  select(id, ends_with("coleop"), ends_with("zeale"))

# subset databases that summarize by order, family, genera, species-level IDs
art_order_reads <- art_full %>%
  filter(order != "NA" & is.na(order)==F) %>%
  group_by(primer, order) %>%
  summarise(across(starts_with("sample"), sum),
            .groups = "drop") %>%
  pivot_wider(names_from = primer, values_from = starts_with("sample")) %>%
  select(order, ends_with("coleop"), ends_with("zeale"))

art_order_motus <- art_full %>%
  filter(order != "NA" & is.na(order)==F) %>%
  group_by(primer, order) %>%
  summarise(across(starts_with("sample"), function(x) sum(x>0)),
            .groups = "drop") %>%
  pivot_wider(names_from = primer, values_from = starts_with("sample")) %>%
  select(order, ends_with("coleop"), ends_with("zeale"))

art_fam_reads <- art_full %>%
  filter(family != "NA" & is.na(family)==F) %>%
  group_by(primer, family) %>%
  summarise(across(starts_with("sample"), sum),
            .groups = "drop") %>%
  pivot_wider(names_from = primer, values_from = starts_with("sample")) %>%
  select(family, ends_with("coleop"), ends_with("zeale"))

art_fam_motus <- art_full %>%
  filter(family != "NA" & is.na(family)==F) %>%
  group_by(primer, family) %>%
  summarise(across(starts_with("sample"), function(x) sum(x>0)),
            .groups = "drop") %>%
  pivot_wider(names_from = primer, values_from = starts_with("sample")) %>%
  select(family, ends_with("coleop"), ends_with("zeale"))

art_gen_reads <- art_full %>%
  filter(genus != "NA" & is.na(genus)==F) %>%
  group_by(primer, genus) %>%
  summarise(across(starts_with("sample"), sum),
            .groups = "drop") %>%
  pivot_wider(names_from = primer, values_from = starts_with("sample")) %>%
  select(genus, ends_with("coleop"), ends_with("zeale"))

art_gen_motus <- art_full %>%
  filter(genus != "NA" & is.na(genus)==F) %>%
  group_by(primer, genus) %>%
  summarise(across(starts_with("sample"), function(x) sum(x>0)),
            .groups = "drop") %>%
  pivot_wider(names_from = primer, values_from = starts_with("sample")) %>%
  select(genus, ends_with("coleop"), ends_with("zeale"))

art_sp_reads <- art_full %>%
  filter(species != "NA" & is.na(species)==F) %>% 
  group_by(primer, species) %>%
  summarise(across(starts_with("sample"), sum),
            .groups = "drop") %>%
  pivot_wider(names_from = primer, values_from = starts_with("sample")) %>%
  select(species, ends_with("coleop"), ends_with("zeale"))

#### Format data for accumulation curve analyses

# create data matrices

# spray samples - zeale - motu level
sprayZ_motu_mat <- art_motu_reads %>%
  filter(grepl(id, pattern = "Z")) %>%
  select(contains("CS"), contains("DS")) %>%
  select(contains("zeale")) %>%
  select(-sample.3CS_zeale) %>% # a sample with no reads for the Zeale primer
  t()
sprayZ_motu_mat[is.na(sprayZ_motu_mat)] <- 0 
sprayZ_motu_mat[sprayZ_motu_mat>0] <- 1 
colnames(sprayZ_motu_mat) <- filter(art_motu_reads, grepl(id, pattern = "Z"))$id

# roller samples - zeale - motu level
rollZ_motu_mat <- art_motu_reads %>%
  filter(grepl(id, pattern = "Z")) %>%
  select(contains("CR"), contains("DR")) %>%
  select(contains("zeale")) %>%
  t()
rollZ_motu_mat[is.na(rollZ_motu_mat)] <- 0 
rollZ_motu_mat[rollZ_motu_mat>0] <- 1 
colnames(rollZ_motu_mat) <- filter(art_motu_reads, grepl(id, pattern = "Z"))$id

# both - zeale - sp level
zeromat.motu.Z <- matrix(0,1,650)
sprayZ_motu_mat2 <- rbind(sprayZ_motu_mat[1:2,],
                         zeromat.motu.Z,
                         sprayZ_motu_mat[3:9,],
                         zeromat.motu.Z,
                         sprayZ_motu_mat[10,],
                         zeromat.motu.Z,
                         sprayZ_motu_mat[11:15,],
                         zeromat.motu.Z,
                         sprayZ_motu_mat[16,])

bothZ_motu_mat <- sprayZ_motu_mat2 + rollZ_motu_mat
bothZ_motu_mat[bothZ_motu_mat>0] <- 1

# spray samples - coleop - motu level
sprayC_motu_mat <- art_motu_reads %>%
  filter(grepl(id, pattern = "C")) %>%
  select(contains("CS"), contains("DS")) %>%
  select(contains("coleop")) %>%
  t()
sprayC_motu_mat[is.na(sprayC_motu_mat)] <- 0 
sprayC_motu_mat[sprayC_motu_mat>0] <- 1 
colnames(sprayC_motu_mat) <- filter(art_motu_reads, grepl(id, pattern = "C"))$id

# roller samples - coleop - motu level
rollC_motu_mat <- art_motu_reads %>%
  filter(grepl(id, pattern = "C")) %>%
  select(contains("CR"), contains("DR")) %>%
  select(contains("coleop")) %>%
  t()
rollC_motu_mat[is.na(rollC_motu_mat)] <- 0 
rollC_motu_mat[rollC_motu_mat>0] <- 1 
colnames(rollC_motu_mat) <- filter(art_motu_reads, grepl(id, pattern = "C"))$id

# both - coleop - motu level
zeromat.motu.C <- matrix(0,1,1077)
sprayC_motu_mat2 <- rbind(sprayC_motu_mat[1:10,],
                         zeromat.motu.C,
                         sprayC_motu_mat[11,],
                         zeromat.motu.C,
                         sprayC_motu_mat[12:16,],
                         zeromat.motu.C,
                         sprayC_motu_mat[17,])

bothC_motu_mat <- sprayC_motu_mat2 + rollC_motu_mat
bothC_motu_mat[bothC_motu_mat>0] <- 1

# spray samples - zeale - sp level
sprayZ_sp_mat <- art_sp_reads %>%
  select(contains("CS"), contains("DS")) %>%
  select(contains("zeale")) %>%
  select(-sample.3CS_zeale) %>% # a sample with no reads for the Zeale primer
  t()
sprayZ_sp_mat[is.na(sprayZ_sp_mat)] <- 0 
sprayZ_sp_mat[sprayZ_sp_mat>0] <- 1 
colnames(sprayZ_sp_mat) <- art_sp_reads$species

# roller samples - zeale - sp level
rollZ_sp_mat <- art_sp_reads %>%
  select(contains("CR"), contains("DR")) %>%
  select(contains("zeale")) %>%
  t()
rollZ_sp_mat[is.na(rollZ_sp_mat)] <- 0 
rollZ_sp_mat[rollZ_sp_mat>0] <- 1 
colnames(rollZ_sp_mat) <- art_sp_reads$species

# both - zeale - sp level
zeromat.sp <- matrix(0,1,292)
sprayZ_sp_mat2 <- rbind(sprayZ_sp_mat[1:2,],
                         zeromat.sp,
                         sprayZ_sp_mat[3:9,],
                         zeromat.sp,
                         sprayZ_sp_mat[10,],
                         zeromat.sp,
                         sprayZ_sp_mat[11:15,],
                         zeromat.sp,
                         sprayZ_sp_mat[16,])

bothZ_sp_mat <- sprayZ_sp_mat2 + rollZ_sp_mat
bothZ_sp_mat[bothZ_sp_mat>0] <- 1

# spray samples - coleop - sp level
sprayC_sp_mat <- art_sp_reads %>%
  select(contains("CS"), contains("DS")) %>%
  select(contains("coleop")) %>%
  t()
sprayC_sp_mat[is.na(sprayC_sp_mat)] <- 0 
sprayC_sp_mat[sprayC_sp_mat>0] <- 1 
colnames(sprayC_sp_mat) <- art_sp_reads$species

# roller samples - coleop - sp level
rollC_sp_mat <- art_sp_reads %>%
  select(contains("CR"), contains("DR")) %>%
  select(contains("coleop")) %>%
  t()
rollC_sp_mat[is.na(rollC_sp_mat)] <- 0 
rollC_sp_mat[rollC_sp_mat>0] <- 1 
colnames(rollC_sp_mat) <- art_sp_reads$species

# both - coleop - sp level
sprayC_sp_mat2 <- rbind(sprayC_sp_mat[1:10,],
                         zeromat.sp,
                         sprayC_sp_mat[11,],
                         zeromat.sp,
                         sprayC_sp_mat[12:16,],
                         zeromat.sp,
                         sprayC_sp_mat[17,])

bothC_sp_mat <- sprayC_sp_mat2 + rollC_sp_mat
bothC_sp_mat[bothC_sp_mat>0] <- 1

# pool all primers and sampling methods - species level
bothCZ_sp_mat <- bothC_sp_mat + bothZ_sp_mat
bothCZ_sp_mat[bothCZ_sp_mat>0] <- 1

# pooled roller samples - species level
roll_sp_mat <- rollC_sp_mat + rollZ_sp_mat
roll_sp_mat[roll_sp_mat>0] <- 1

# pooled spray samples - species level
spray_sp_mat <- sprayC_sp_mat2 + sprayZ_sp_mat2
spray_sp_mat[spray_sp_mat>0] <- 1

# spray samples - zeale - genus level
sprayZ_gen_mat <- art_gen_reads %>%
  select(contains("CS"), contains("DS")) %>%
  select(contains("zeale")) %>%
  select(-sample.3CS_zeale) %>% # a sample with no reads for the Zeale primer
  t()
sprayZ_gen_mat[is.na(sprayZ_gen_mat)] <- 0 
sprayZ_gen_mat[sprayZ_gen_mat>0] <- 1 
colnames(sprayZ_gen_mat) <- art_gen_reads$genus

# roller samples - zeale - genus level
rollZ_gen_mat <- art_gen_reads %>%
  select(contains("CR"), contains("DR")) %>%
  select(contains("zeale")) %>%
  t()
rollZ_gen_mat[is.na(rollZ_gen_mat)] <- 0 
rollZ_gen_mat[rollZ_gen_mat>0] <- 1 
colnames(rollZ_gen_mat) <- art_gen_reads$genus

# both - zeale - genus level
zeromat.gen <- matrix(0,1,353)
sprayZ_gen_mat2 <- rbind(sprayZ_gen_mat[1:2,],
                         zeromat.gen,
                         sprayZ_gen_mat[3:9,],
                         zeromat.gen,
                         sprayZ_gen_mat[10,],
                         zeromat.gen,
                         sprayZ_gen_mat[11:15,],
                         zeromat.gen,
                         sprayZ_gen_mat[16,])

bothZ_gen_mat <- sprayZ_gen_mat2 + rollZ_gen_mat
bothZ_gen_mat[bothZ_gen_mat>0] <- 1

# spray samples - coleop - genus level
sprayC_gen_mat <- art_gen_reads %>%
  select(contains("CS"), contains("DS")) %>%
  select(contains("coleop")) %>%
  t()
sprayC_gen_mat[is.na(sprayC_gen_mat)] <- 0 
sprayC_gen_mat[sprayC_gen_mat>0] <- 1 
colnames(sprayC_gen_mat) <- art_gen_reads$genus

# roller samples - coleop - genus level
rollC_gen_mat <- art_gen_reads %>%
  select(contains("CR"), contains("DR")) %>%
  select(contains("coleop")) %>%
  t()
rollC_gen_mat[is.na(rollC_gen_mat)] <- 0 
rollC_gen_mat[rollC_gen_mat>0] <- 1 
colnames(rollC_gen_mat) <- art_gen_reads$genus

# both - coleop - genus level
sprayC_gen_mat2 <- rbind(sprayC_gen_mat[1:10,],
                         zeromat.gen,
                         sprayC_gen_mat[11,],
                         zeromat.gen,
                         sprayC_gen_mat[12:16,],
                         zeromat.gen,
                         sprayC_gen_mat[17,])

bothC_gen_mat <- sprayC_gen_mat2 + rollC_gen_mat
bothC_gen_mat[bothC_gen_mat>0] <- 1

# pool all primers and sampling methods - genus level
bothCZ_gen_mat <- bothC_gen_mat + bothZ_gen_mat
bothCZ_gen_mat[bothCZ_gen_mat>0] <- 1

# pooled roller samples - genus level
roll_gen_mat <- rollC_gen_mat + rollZ_gen_mat
roll_gen_mat[roll_gen_mat>0] <- 1

# pooled spray samples - genus level
spray_gen_mat <- sprayC_gen_mat2 + sprayZ_gen_mat2
spray_gen_mat[spray_gen_mat>0] <- 1

# spray samples - zeale - family level
sprayZ_fam_mat <- art_fam_reads %>%
  select(contains("CS"), contains("DS")) %>%
  select(contains("zeale")) %>%
  select(-sample.3CS_zeale) %>% # a sample with no reads for the Zeale primer
  t()
sprayZ_fam_mat[is.na(sprayZ_fam_mat)] <- 0 
sprayZ_fam_mat[sprayZ_fam_mat>0] <- 1 
colnames(sprayZ_fam_mat) <- art_fam_reads$family

# roller samples - zeale primer - family level
rollZ_fam_mat <- art_fam_reads %>%
  select(contains("CR"), contains("DR")) %>%
  select(contains("zeale")) %>%
  t()
rollZ_fam_mat[is.na(rollZ_fam_mat)] <- 0 
rollZ_fam_mat[rollZ_fam_mat>0] <- 1 
colnames(rollZ_fam_mat) <- art_fam_reads$family

# both - zeale - family level
zeromat <- matrix(0,1,181)
sprayZ_fam_mat2 <- rbind(sprayZ_fam_mat[1:2,],
                         zeromat,
                         sprayZ_fam_mat[3:9,],
                         zeromat,
                         sprayZ_fam_mat[10,],
                         zeromat,
                         sprayZ_fam_mat[11:15,],
                         zeromat,
                         sprayZ_fam_mat[16,])

bothZ_fam_mat <- sprayZ_fam_mat2 + rollZ_fam_mat
bothZ_fam_mat[bothZ_fam_mat>0] <- 1

# spray samples - coleop - family level
sprayC_fam_mat <- art_fam_reads %>%
  select(contains("CS"), contains("DS")) %>%
  select(contains("coleop")) %>%
  t()
sprayC_fam_mat[is.na(sprayC_fam_mat)] <- 0 
sprayC_fam_mat[sprayC_fam_mat>0] <- 1 
colnames(sprayC_fam_mat) <- art_fam_reads$family

# roller samples - coleop primer - family level
rollC_fam_mat <- art_fam_reads %>%
  select(contains("CR"), contains("DR")) %>%
  select(contains("coleop")) %>%
  t()
rollC_fam_mat[is.na(rollC_fam_mat)] <- 0 
rollC_fam_mat[rollC_fam_mat>0] <- 1 
colnames(rollC_fam_mat) <- art_fam_reads$family

# both - coleop - family level
sprayC_fam_mat2 <- rbind(sprayC_fam_mat[1:10,],
                         zeromat,
                         sprayC_fam_mat[11,],
                         zeromat,
                         sprayC_fam_mat[12:16,],
                         zeromat,
                         sprayC_fam_mat[17,])

bothC_fam_mat <- sprayC_fam_mat2 + rollC_fam_mat
bothC_fam_mat[bothC_fam_mat>0] <- 1

# pool all primers and sampling methods - family level
bothCZ_fam_mat <- bothC_fam_mat + bothZ_fam_mat
bothCZ_fam_mat[bothCZ_fam_mat>0] <- 1

# pooled roller samples - family level
roll_fam_mat <- rollC_fam_mat + rollZ_fam_mat
roll_fam_mat[roll_fam_mat>0] <- 1

# pooled spray samples - family level
spray_fam_mat <- sprayC_fam_mat2 + sprayZ_fam_mat2
spray_fam_mat[spray_fam_mat>0] <- 1

# read counts by unique taxon
readsC = cs3 %>% group_by(phylum, class, order, family, genus, species) %>% summarize(count = sum(count))
readsZ = zs3 %>% group_by(phylum, class, order, family, genus, species) %>% summarize(count = sum(count))

```
# Create Figure 1
Create Figure 1 in Allen et al. (2023). MOTU counts by order, primer set, and understory foliage vs. bark.
```{r}
#################
#### Plotting abundance of each order by method (bar charts)
#### (no. MOTUs)
#################

### Coleop ###

# format data for spray samples (coleop)
ordersCspray <- art_full %>%
  filter(!is.na(order), primer == "coleop") %>%
  select(class, order, family, genus, species, contains(c("CS", "DS"))) %>%
  rowwise() %>% 
  mutate(sample.sum = sum(c_across(starts_with("sample")), na.rm = T)) %>%
  ungroup() %>%
  mutate(idsp = case_when(!is.na(species) & sample.sum > 0 ~ 1,
                          TRUE ~ 0),
         idgen = case_when(!is.na(genus) & is.na(species) & sample.sum > 0 ~ 1,
                           TRUE ~ 0),
         idfam = case_when((!is.na(family) & is.na(genus) & sample.sum > 0) ~ 1,
                           TRUE ~ 0),
         idord = case_when((!is.na(order) & is.na(family) & sample.sum > 0) ~ 1,
                           TRUE ~ 0)) %>%
  select(order, idsp, idgen, idfam, idord, sample.sum) %>%
  group_by(order) %>%
  summarize(motus = sum(sample.sum > 0),
            ord = sum(idord),
            fam = sum(idfam),
            gen = sum(idgen),
            sp = sum(idsp),
            .groups = "drop") %>%
  right_join(select(art_order_reads, order)) %>%
  tidyr::replace_na(list(motus = 0, ord = 0, fam = 0, gen = 0, sp = 0))%>%
  left_join(distinct(select(lineage_lookup, class, order))) %>%
  arrange(class, order) %>%
  mutate(primer = "Understory (16S)")

# format data for roller samples (coleop)
ordersCroll <- art_full %>%
  filter(!is.na(order), primer == "coleop") %>%
  select(class, order, family, genus, species, contains(c("CR", "DR"))) %>%
  rowwise() %>% 
  mutate(sample.sum = sum(c_across(starts_with("sample")), na.rm = T)) %>%
  ungroup() %>%
  mutate(idsp = case_when(!is.na(species) & sample.sum > 0 ~ 1,
                          TRUE ~ 0),
         idgen = case_when(!is.na(genus) & is.na(species) & sample.sum > 0 ~ 1,
                           TRUE ~ 0),
         idfam = case_when((!is.na(family) & is.na(genus) & sample.sum > 0) ~ 1,
                           TRUE ~ 0),
         idord = case_when((!is.na(order) & is.na(family) & sample.sum > 0) ~ 1,
                           TRUE ~ 0)) %>%
  select(order, idsp, idgen, idfam, idord, sample.sum) %>%
  group_by(order) %>%
  summarize(motus = sum(sample.sum > 0),
            ord = sum(idord),
            fam = sum(idfam),
            gen = sum(idgen),
            sp = sum(idsp),
            .groups = "drop") %>%
  right_join(select(art_order_reads, order)) %>%
  tidyr::replace_na(list(motus = 0, ord = 0, fam = 0, gen = 0, sp = 0))%>%
  left_join(distinct(select(lineage_lookup, class, order))) %>%
  arrange(class, order) %>%
  mutate(primer = "Bark (16S)")

### Zeale ###

# format data for spray samples (zeale)
ordersZspray <- art_full %>%
  filter(!is.na(order), primer == "zeale") %>%
  select(class, order, family, genus, species, contains(c("CS", "DS"))) %>%
  rowwise() %>% 
  mutate(sample.sum = sum(c_across(starts_with("sample")), na.rm = T)) %>%
  ungroup() %>%
  mutate(idsp = case_when(!is.na(species) & sample.sum > 0 ~ 1,
                          TRUE ~ 0),
         idgen = case_when(!is.na(genus) & is.na(species) & sample.sum > 0 ~ 1,
                           TRUE ~ 0),
         idfam = case_when((!is.na(family) & is.na(genus) & sample.sum > 0) ~ 1,
                           TRUE ~ 0),
         idord = case_when((!is.na(order) & is.na(family) & sample.sum > 0) ~ 1,
                           TRUE ~ 0)) %>%
  select(order, idsp, idgen, idfam, idord, sample.sum) %>%
  group_by(order) %>%
  summarize(motus = sum(sample.sum > 0),
            ord = sum(idord),
            fam = sum(idfam),
            gen = sum(idgen),
            sp = sum(idsp),
            .groups = "drop") %>%
  right_join(select(art_order_reads, order)) %>%
  tidyr::replace_na(list(motus = 0, ord = 0, fam = 0, gen = 0, sp = 0))%>%
  left_join(distinct(select(lineage_lookup, class, order))) %>%
  arrange(class, order) %>%
  mutate(primer = "Understory (COI)")

# format data for roll samples (zeale)
ordersZroll <- art_full %>%
  filter(!is.na(order), primer == "zeale") %>%
  select(class, order, family, genus, species, contains(c("CR", "DR"))) %>%
  rowwise() %>% 
  mutate(sample.sum = sum(c_across(starts_with("sample")), na.rm = T)) %>%
  ungroup() %>%
  mutate(idsp = case_when(!is.na(species) & sample.sum > 0 ~ 1,
                          TRUE ~ 0),
         idgen = case_when(!is.na(genus) & is.na(species) & sample.sum > 0 ~ 1,
                           TRUE ~ 0),
         idfam = case_when((!is.na(family) & is.na(genus) & sample.sum > 0) ~ 1,
                           TRUE ~ 0),
         idord = case_when((!is.na(order) & is.na(family) & sample.sum > 0) ~ 1,
                           TRUE ~ 0)) %>%
  select(order, idsp, idgen, idfam, idord, sample.sum) %>%
  group_by(order) %>%
  summarize(motus = sum(sample.sum > 0),
            ord = sum(idord),
            fam = sum(idfam),
            gen = sum(idgen),
            sp = sum(idsp),
            .groups = "drop") %>%
  right_join(select(art_order_reads, order)) %>%
  tidyr::replace_na(list(motus = 0, ord = 0, fam = 0, gen = 0, sp = 0))%>%
  left_join(distinct(select(lineage_lookup, class, order))) %>%
  arrange(class, order) %>%
  mutate(primer = "Bark (COI)")

# combine MOTU count datasets into one for plotting

orders_df2 <- ordersCspray %>%
  bind_rows(ordersZspray) %>%
  bind_rows(ordersCroll) %>%
  bind_rows(ordersZroll) %>%
  mutate(engname = case_when(order == "Mesostigmata" ~ "Mites",
                          order == "Sarcoptiformes" ~ "Mites",
                          order == "Trombidiformes" ~ "Mites",
                          class == "Collembola" ~ "Springtails",
                          class == "Chilopoda" ~ "Centipedes",
                          order == "Araneae" ~ "Spiders",
                          order == "Opiliones" ~ "Harvestmen",
                          order == "Blattodea" ~ "Cockroaches",
                          order == "Coleoptera" ~ "Beetles",
                          order == "Dermaptera" ~ "Earwigs",
                          order == "Diptera" ~ "Flies",
                          order == "Ephemeroptera" ~ "Mayflies",
                          order == "Hemiptera" ~ "True bugs",
                          order == "Hymenoptera" ~ "Bees, wasps, & ants",
                          order == "Lepidoptera" ~ "Butterflies & moths",
                          order == "Mantodea" ~ "Mantids",
                          order == "Megaloptera" ~ "Dobsonflies",
                          order == "Cyclopoida" ~ "Copepods",
                          order == "Neuroptera" ~ "Lacewings",
                          order == "Odonata" ~ "Dragonflies & damselflies",
                          order == "Orthoptera" ~ "Crickets & grasshoppers",
                          order == "Phasmatodea" ~ "Walking sticks",
                          order == "Phthiraptera" ~ "Thrips",
                          order == "Psocoptera" ~ "Bark lice",
                          order == "Siphonaptera" ~ "Fleas",
                          order == "Trichoptera" ~ "Caddisflies"
                          )) %>%
  group_by(class, primer, engname) %>%
  summarize(Order = sum(ord, na.rm = T),
            Family = sum(fam, na.rm = T),
            Genus = sum(gen, na.rm = T),
            Species = sum(sp, na.rm = T),
            .groups = "drop") %>%
  pivot_longer(Order:Species, names_to = "idlevel", values_to = "motus") %>%
  group_by(class, engname) %>%
  mutate(summotu = sum(motus)) %>%
  ungroup() %>%
  arrange(primer, class, engname, idlevel) %>%
  mutate(taxorder = case_when(idlevel == "Species" ~ 1,
                              idlevel == "Genus" ~ 2,
                              idlevel == "Family" ~ 3,
                              idlevel == "Order" ~ 4),
         primorder = case_when(primer == "Understory (COI)" ~ 1,
                              primer == "Bark (COI)" ~ 2,
                              primer == "Understory (16S)" ~ 3,
                              primer == "Bark (16S)" ~ 4)) %>%
  mutate(engname = fct_reorder(.f = engname, .x = summotu, .desc = F),
         idlevel = fct_reorder(.f = idlevel, .x = taxorder, .desc = T),
         primer = fct_reorder(.f = primer, .x = primorder, .desc = T))

# make plot for less-diverse taxa
library(forcats)
(
ldiv_plot <- 
orders_df2 %>%
    mutate(group = case_when(summotu >= 50 ~ "More",
                             TRUE ~ "Less")) %>%
    filter(group == "Less") %>%
  ggplot() +
  geom_col(aes(y = engname, x = motus,
               fill = idlevel)) +
  facet_wrap(~ primer, ncol = 4) +
  # scale_fill_manual(values = wes_palettes$Zissou1) +
  # scale_color_manual(values = wes_palettes$Zissou1) +
  scale_fill_manual(values = c("gray10", "gray30", "gray50", "gray70")) +
  scale_color_manual(values = c("gray10", "gray30", "gray50", "gray70")) +
  theme_bw() +
  theme(text = element_text(size = 14),
        axis.text.y = element_text(size = 11), 
        legend.position = c(.89,.22),
        legend.background = element_blank(),
        legend.box.background = element_rect(colour = "black")
        ) +
  theme(legend.key.size = unit(.45, 'cm'), #change legend key size
        legend.key.height = unit(.45, 'cm'), #change legend key height
        legend.key.width = unit(.45, 'cm'), #change legend key width
        legend.title = element_text(size=12), #change legend title font size
        legend.text = element_text(size=12)) +
  labs(x = "Number of MOTUs",
       y = "",
       fill = "Lowest \ntaxonomic \nID") #+
  #  guides(fill = "none")
)

# make plot for more-diverse taxa
(
hdiv_plot <- 
orders_df2 %>%
    mutate(group = case_when(summotu >= 50 ~ "More",
                             TRUE ~ "Less")) %>%
    filter(group == "More") %>%
  ggplot() +
  geom_col(aes(y = engname, x = motus,
               fill = idlevel)) +
  facet_wrap(~ primer, ncol = 4) +
  # scale_fill_manual(values = wes_palettes$Darjeeling1) +
  # scale_color_manual(values = wes_palettes$Darjeeling1) +
  scale_fill_manual(values = c("gray10", "gray30", "gray50", "gray70")) +
  scale_color_manual(values = c("gray10", "gray30", "gray50", "gray70")) +
  theme_bw() +
  theme(text = element_text(size = 14),
        axis.text.y = element_text(size = 11)#, legend.position = c(.89,.24)
        ) +
  labs(x = "",
       y = "",
       fill = "Lowest \ntaxonomic \nID") +
    guides(fill = "none")
)

library(patchwork)
hdiv_plot / ldiv_plot

# ggsave("figures/Figure1.png", height = 9, width = 7.5, dpi = 400)
```
# Create Table 1
Create Table 1 in Allen et al. (2023). Mean reads by order and sample substrate.
```{r}
master <- art_full  %>%
  mutate(engname = case_when(order == "Mesostigmata" ~ "Mites",
                          order == "Sarcoptiformes" ~ "Mites",
                          order == "Trombidiformes" ~ "Mites",
                          class == "Collembola" ~ "Springtails",
                          class == "Chilopoda" ~ "Centipedes",
                          order == "Araneae" ~ "Spiders",
                          order == "Opiliones" ~ "Harvestmen",
                          order == "Blattodea" ~ "Cockroaches",
                          order == "Coleoptera" ~ "Beetles",
                          order == "Dermaptera" ~ "Earwigs",
                          order == "Diptera" ~ "Flies",
                          order == "Ephemeroptera" ~ "Mayflies",
                          order == "Hemiptera" ~ "True bugs",
                          order == "Hymenoptera" ~ "Bees, wasps, & ants",
                          order == "Lepidoptera" ~ "Butterflies & moths",
                          order == "Mantodea" ~ "Mantids",
                          order == "Megaloptera" ~ "Dobsonflies",
                          order == "Cyclopoida" ~ "Copepods",
                          order == "Neuroptera" ~ "Lacewings",
                          order == "Odonata" ~ "Dragonflies & damselflies",
                          order == "Orthoptera" ~ "Crickets & grasshoppers",
                          order == "Phasmatodea" ~ "Walking sticks",
                          order == "Phthiraptera" ~ "Thrips",
                          order == "Psocoptera" ~ "Bark lice",
                          order == "Siphonaptera" ~ "Fleas",
                          order == "Trichoptera" ~ "Caddisflies",
                          (is.na(order) & is.na(class) & phylum == "Arthropoda") ~ "Arthropod sp.",
                          (is.na(order) & !is.na(class) ~ paste0(class, " sp."))
                          ))

# format data for spray samples
orderreads_spray <- master %>%
  group_by(engname, primer) %>%
  summarize(across(contains(c("CS", "DS")), sum),
            .groups = "drop") %>%
  rowwise() %>% 
  mutate(sample.sum = sum(c_across(contains(c("CS", "DS"))), na.rm = T),
         sample.mean = mean(c_across(contains(c("CS", "DS"))), na.rm = T),
         sample.sd = sd(c_across(contains(c("CS", "DS"))), na.rm = T),
         sample.n = sum(!is.na(c_across(contains(c("CS", "DS")))), na.rm = T),
         sample.pres = sum((c_across(contains(c("CS", "DS")))>0), na.rm = T)) %>%
  ungroup() %>%
  mutate(sample.se = sample.sd / sqrt(sample.n),
         sample.pct = 100*(sample.pres / sample.n)) %>%
  right_join(expand.grid(engname = unique(master$engname), primer = c("coleop", "zeale"))) %>%
  filter(!is.na(engname)) %>%
  select(engname, primer, contains(c("mean", ".se", "pct"))) %>%
  tidyr::replace_na(list(sample.mean = 0, sample.pct = 0, sample.se = 0)) %>%
  mutate(US_meanse = paste0(round(sample.mean,0), " (", round(sample.se, 0), ")"), # \u00B1 
         US_pct = round(sample.pct,1)) %>%
  rename(US_mean = sample.mean) %>%
  left_join(distinct(select(master, class, engname))) %>%
  arrange(primer, class, desc(US_mean)) %>%
  pivot_wider(id_cols = c(engname), names_from = primer, values_from = c(US_meanse, US_pct, US_mean)) %>%
  left_join(distinct(select(master, class, engname)))

# format data for roller samples
orderreads_roll <- master %>%
  filter(!is.na(order)) %>%
  group_by(engname, primer) %>%
  summarize(across(contains(c("CR", "DR")), sum),
            .groups = "drop") %>%
  rowwise() %>% 
  mutate(sample.sum = sum(c_across(contains(c("CR", "DR"))), na.rm = T),
         sample.mean = mean(c_across(contains(c("CR", "DR"))), na.rm = T),
         sample.sd = sd(c_across(contains(c("CR", "DR"))), na.rm = T),
         sample.n = sum(!is.na(c_across(contains(c("CR", "DR")))), na.rm = T),
         sample.pres = sum((c_across(contains(c("CR", "DR")))>0), na.rm = T)) %>%
  ungroup() %>%
  mutate(sample.se = sample.sd / sqrt(sample.n),
         sample.pct = 100*(sample.pres / sample.n)) %>%
  right_join(expand.grid(engname = unique(master$engname), primer = c("coleop", "zeale"))) %>%
  filter(!is.na(engname)) %>%
  select(engname, primer, contains(c("mean", ".se", "pct"))) %>%
  tidyr::replace_na(list(sample.mean = 0, sample.pct = 0, sample.se = 0)) %>%
  mutate(B_meanse = paste0(round(sample.mean,0), " (", round(sample.se, 0), ")"), # \u00B1 
         B_pct = round(sample.pct,1)) %>%
  rename(B_mean = sample.mean) %>%
  left_join(distinct(select(master, class, engname))) %>%
  arrange(primer, class, desc(B_mean)) %>%
  pivot_wider(id_cols = c(engname), names_from = primer, values_from = c(B_meanse, B_pct, B_mean)) %>%
  left_join(distinct(select(master, class, engname)))

# combine read count data sets into one for table

ordereads_df <- orderreads_spray %>%
  left_join(orderreads_roll) %>%
  rowwise() %>%
  mutate(sort = mean(c_across(contains(c("B_mean_", "US_mean_"))))) %>%
  ungroup() %>%
  arrange(class, desc(sort)) %>%
  select(class, engname, B_meanse_zeale, B_pct_zeale, US_meanse_zeale, US_pct_zeale,
         B_meanse_coleop, B_pct_coleop, US_meanse_coleop, US_pct_coleop)
  
library(knitr) # knitr_1.43 
library(kableExtra) # kableExtra_1.3.4
library(magrittr) # magrittr_2.0.3

kable(ordereads_df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  cat(., file = "figures/Table1.html")
```

# Create Figure 2
Create Figure 2 in Allen et al. (2023). Compute accumulation curves using iNEXT.
```{r}
# format data for iNEXT
accum.data.motu <- list('Understory (COI)' = t(sprayZ_motu_mat),
                 'Bark (COI)' = t(rollZ_motu_mat),
                 'Both substrates (COI)' = t(bothZ_motu_mat),
                 'Understory (16S)' = t(sprayC_motu_mat),
                 'Bark (16S)' = t(rollC_motu_mat),
                 'Both substrates (16S)' = t(bothC_motu_mat))

accum.data.motu.primer <- list('Primer COI' = t(bothZ_motu_mat),
                 'Primer 16S' = t(bothC_motu_mat))

accum.data.motu.sub <- list('Understory (COI)' = t(sprayZ_motu_mat),
                 'Bark (COI)' = t(rollZ_motu_mat),
                 'Understory (16S)' = t(sprayC_motu_mat),
                 'Bark (16S)' = t(rollC_motu_mat))

accum.data.sp.primer <- list(
                 'Primer COI' = t(bothZ_sp_mat),
                 'Primer 16S' = t(bothC_sp_mat),
                 'All pooled' = t(bothCZ_sp_mat))

accum.data.sp.sub <- list(
                 'Bark' = t(roll_sp_mat),
                 'Understory' = t(spray_sp_mat),
                 'All pooled' = t(bothCZ_sp_mat))

accum.data.gen.primer <- list(
                 'Primer COI' = t(bothZ_gen_mat),
                 'Primer 16S' = t(bothC_gen_mat),
                 'All pooled' = t(bothCZ_gen_mat))

accum.data.gen.sub <- list(
                 'Bark' = t(roll_gen_mat),
                 'Understory' = t(spray_gen_mat),
                 'All pooled' = t(bothCZ_gen_mat))

accum.data.fam.primer <- list(
                 'Primer COI' = t(bothZ_fam_mat),
                 'Primer 16S' = t(bothC_fam_mat),
                 'All pooled' = t(bothCZ_fam_mat))

accum.data.fam.sub <- list(
                 'Bark' = t(roll_fam_mat),
                 'Understory' = t(spray_fam_mat),
                 'All pooled' = t(bothCZ_fam_mat))

# run accumulation curve analyses in iNEXT
accum.motu.primer <- iNEXT(accum.data.motu.primer, q=0, datatype="incidence_raw", endpoint = 100)
# accum.motu.primer$iNextEst$coverage_based

accum.motu.sub <- iNEXT(accum.data.motu.sub, q=0, datatype="incidence_raw", endpoint = 100)
# accum.motu$iNextEst$coverage_based

accum.sp.primer <- iNEXT(accum.data.sp.primer, q=0, datatype="incidence_raw", endpoint = 100)
# accum.sp.primer$iNextEst$coverage_based

accum.sp.sub <- iNEXT(accum.data.sp.sub, q=0, datatype="incidence_raw", endpoint = 100)
# accum.sp.sub$iNextEst$coverage_based

accum.gen.primer <- iNEXT(accum.data.gen.primer, q=0, datatype="incidence_raw", endpoint = 100)
# accum.gen.primer$iNextEst$coverage_based

accum.gen.sub <- iNEXT(accum.data.gen.sub, q=0, datatype="incidence_raw", endpoint = 100)
# accum.gen.sub$iNextEst$coverage_based

accum.fam.primer <- iNEXT(accum.data.fam.primer, q=0, datatype="incidence_raw", endpoint = 100)
# accum.fam.primer$iNextEst$coverage_based

accum.fam.sub <- iNEXT(accum.data.fam.sub, q=0, datatype="incidence_raw", endpoint = 100)
# accum.fam.sub$iNextEst$coverage_based

# create individual plots for motu, sp, genus, family

(motuplot.primer <-
ggiNEXT(accum.motu.primer, type=1, se=TRUE, facet.var="None", color.var="Assemblage", grey=FALSE) +
  scale_x_continuous(limits = c(0,80)) +
  scale_y_continuous(limits = c(0,2100)) +
  labs(y = "No. MOTUs",
       x = "No. samples") +
    theme_bw() +
    scale_color_manual(values = c("darkgreen", "darkorange3")) +
    scale_fill_manual(values = c("darkgreen", "darkorange3")) +
  theme(text = element_text(size = 14),
        legend.position = c(0.65, 0.1),
        legend.background = element_rect(fill = "transparent")) +
  guides(linetype = "none") +
  theme(legend.key.size = unit(.07, 'cm'), #change legend key size
        legend.key.height = unit(.07, 'cm'), #change legend key height
        legend.key.width = unit(.07, 'cm'), #change legend key width
        legend.title = element_text(size=7), #change legend title font size
        legend.text = element_text(size=7)) #change legend text font size
)

(motuplot.sub <-
ggiNEXT(accum.motu.sub, type=1, se=TRUE, facet.var="None", color.var="Assemblage", grey=FALSE) +
  scale_x_continuous(limits = c(0,80)) +
  scale_y_continuous(limits = c(0,2000)) +
  labs(y = "",
       x = "No. samples") +
    theme_bw() +
    scale_color_manual(values = c("darkblue", "steelblue", "darkred", "indianred3")) +
    scale_fill_manual(values = c("darkblue", "steelblue", "darkred", "indianred3")) +
  theme(text = element_text(size = 14),
        legend.position = c(0.65, 0.1),
        legend.background = element_rect(fill = "transparent")) +
  guides(linetype = "none") +
  theme(legend.key.size = unit(.07, 'cm'), #change legend key size
        legend.key.height = unit(.07, 'cm'), #change legend key height
        legend.key.width = unit(.07, 'cm'), #change legend key width
        legend.title = element_text(size=7), #change legend title font size
        legend.text = element_text(size=7)) #change legend text font size
)

(spplot.primer <-
ggiNEXT(accum.sp.primer, type=1, se=TRUE, facet.var="None", color.var="Assemblage", grey=FALSE) +
  scale_x_continuous(limits = c(0,80)) +
  labs(y = "No. species identified",
       x = "No. samples") +
    theme_bw() +
    scale_color_manual(values = c("black", "darkgreen", "darkorange3")) +
    scale_fill_manual(values = c("black", "darkgreen", "darkorange3")) +
  theme(text = element_text(size = 14),
        legend.position = c(0.65, 0.1),
        legend.background = element_rect(fill = "transparent")) +
  guides(linetype = "none") +
  theme(legend.key.size = unit(.07, 'cm'), #change legend key size
        legend.key.height = unit(.07, 'cm'), #change legend key height
        legend.key.width = unit(.07, 'cm'), #change legend key width
        legend.title = element_text(size=7), #change legend title font size
        legend.text = element_text(size=7)) #change legend text font size
)

(spplot.sub <-
ggiNEXT(accum.sp.sub, type=1, se=TRUE, facet.var="None", color.var="Assemblage", grey=FALSE) +
  scale_x_continuous(limits = c(0,80)) +
  labs(y = "",
       x = "No. samples") +
    theme_bw() +
    scale_color_manual(values = c("black", "steelblue", "indianred3")) +
    scale_fill_manual(values = c("black", "steelblue", "indianred3")) +
  theme(text = element_text(size = 14),
        legend.position = c(0.65, 0.1),
        legend.background = element_rect(fill = "transparent")) +
  guides(linetype = "none") +
  theme(legend.key.size = unit(.07, 'cm'), #change legend key size
        # legend.key.height = unit(.07, 'cm'), #change legend key height
        # legend.key.width = unit(.07, 'cm'), #change legend key width
        legend.title = element_text(size=7), #change legend title font size
        legend.text = element_text(size=7)) #change legend text font size
)

(genplot.primer <-
ggiNEXT(accum.gen.primer, type=1, se=TRUE, facet.var="None", color.var="Assemblage", grey=FALSE) +
  scale_x_continuous(limits = c(0,80)) +
  labs(y = "No. genera identified",
       x = "No. samples") +
    theme_bw() +
    scale_color_manual(values = c("black", "darkgreen", "darkorange3")) +
    scale_fill_manual(values = c("black", "darkgreen", "darkorange3")) +
  theme(text = element_text(size = 14),
        legend.position = c(0.65, 0.1),
        legend.background = element_rect(fill = "transparent")) +
  guides(linetype = "none") +
  theme(legend.key.size = unit(.07, 'cm'), #change legend key size
        legend.key.height = unit(.07, 'cm'), #change legend key height
        legend.key.width = unit(.07, 'cm'), #change legend key width
        legend.title = element_text(size=7), #change legend title font size
        legend.text = element_text(size=7)) #change legend text font size
)

(genplot.sub <-
ggiNEXT(accum.gen.sub, type=1, se=TRUE, facet.var="None", color.var="Assemblage", grey=FALSE) +
  scale_x_continuous(limits = c(0,80)) +
  labs(y = "",
       x = "No. samples") +
    theme_bw() +
    scale_color_manual(values = c("black", "steelblue", "indianred3")) +
    scale_fill_manual(values = c("black", "steelblue", "indianred3")) +
  theme(text = element_text(size = 14),
        legend.position = c(0.65, 0.1),
        legend.background = element_rect(fill = "transparent")) +
  guides(linetype = "none") +
  theme(legend.key.size = unit(.07, 'cm'), #change legend key size
        # legend.key.height = unit(.07, 'cm'), #change legend key height
        # legend.key.width = unit(.07, 'cm'), #change legend key width
        legend.title = element_text(size=7), #change legend title font size
        legend.text = element_text(size=7)) #change legend text font size
)

(famplot.primer <-
ggiNEXT(accum.fam.primer, type=1, se=TRUE, facet.var="None", color.var="Assemblage", grey=FALSE) +
  scale_x_continuous(limits = c(0,80)) +
  labs(y = "No. families identified",
       x = "No. samples") +
    theme_bw() +
    scale_color_manual(values = c("black", "darkgreen", "darkorange3")) +
    scale_fill_manual(values = c("black", "darkgreen", "darkorange3")) +
  theme(text = element_text(size = 14),
        legend.position = c(0.65, 0.1),
        legend.background = element_rect(fill = "transparent")) +
  guides(linetype = "none") +
  theme(legend.key.size = unit(.07, 'cm'), #change legend key size
        legend.key.height = unit(.07, 'cm'), #change legend key height
        legend.key.width = unit(.07, 'cm'), #change legend key width
        legend.title = element_text(size=7), #change legend title font size
        legend.text = element_text(size=7)) #change legend text font size
)

(famplot.sub <-
ggiNEXT(accum.fam.sub, type=1, se=TRUE, facet.var="None", color.var="Assemblage", grey=FALSE) +
  scale_x_continuous(limits = c(0,80)) +
  labs(y = "",
       x = "No. samples") +
    theme_bw() +
    scale_color_manual(values = c("black", "steelblue", "indianred3")) +
    scale_fill_manual(values = c("black", "steelblue", "indianred3")) +
  theme(text = element_text(size = 14),
        legend.position = c(0.65, 0.1),
        legend.background = element_rect(fill = "transparent")) +
  guides(linetype = "none") +
  theme(legend.key.size = unit(.07, 'cm'), #change legend key size
        # legend.key.height = unit(.07, 'cm'), #change legend key height
        # legend.key.width = unit(.07, 'cm'), #change legend key width
        legend.title = element_text(size=7), #change legend title font size
        legend.text = element_text(size=7)) #change legend text font size
)

# arrange and save all as one plot using patchwork
library(patchwork)
motuplot.primer + motuplot.sub + plot_spacer() + spplot.primer + spplot.sub + 
  genplot.primer + genplot.sub + plot_spacer() + famplot.primer + famplot.sub + 
  plot_layout(ncol = 5, widths = c(0.5, 0.5, 0.15, 0.5, 0.5, 0.5, 0.5, 0.15, 0.5, 0.5))
# ggsave("figures/Figure2.png", dpi = 400, width = 11, height = 10)
```
# Create Figure 3
Create Figure 3 in Allen et al. 2023. Model and plot diversity, richness, and evenness by substrate, tree type, and primer set. Read data rarefied to 20000 reads (drops 6 samples).
```{r}
# MOTU rarefied species richness by sample and primer set
names(art_full) # to get column numbers of matrix
data_cols <- 9:45
coleop_rare_mat <- art_full %>%
  filter(primer == "coleop") %>%
  select(all_of(data_cols)) %>%
  as.matrix() %>%
  t() %>%
  as.data.frame() %>%
  tibble::rownames_to_column() %>%
  rowwise() %>% 
  mutate(sample.sum = sum(c_across(starts_with("V")), na.rm = T)) %>%
  as.data.frame()
rownames(coleop_rare_mat) <- coleop_rare_mat$rowname

zeale_rare_mat <- art_full %>%
  filter(primer == "zeale") %>%
  select(all_of(data_cols)) %>%
  as.matrix() %>%
  t() %>%
  as.data.frame() %>%
  tibble::rownames_to_column() %>%
  rowwise() %>% 
  mutate(sample.sum = sum(c_across(starts_with("V")), na.rm = T)) %>%
  as.data.frame()
rownames(zeale_rare_mat) <- zeale_rare_mat$rowname

# get list of samples with low read depth (< 20000)
lowreadsampsC20 <- coleop_rare_mat %>%
  filter(sample.sum < 20000); (lowreadsampsC20 <- lowreadsampsC20$rowname)

lowreadsampsZ20 <- zeale_rare_mat %>%
  filter(sample.sum < 20000); (lowreadsampsZ20 <- lowreadsampsZ20$rowname)

# create versions of the dataset with only samples reaching threshold read depths

# 20000 threshold removes 6 samples ...
coleop_rare_mat20 <- coleop_rare_mat %>%
  filter(rowname %notin% unique(c(lowreadsampsC20, lowreadsampsZ20))) %>%
  select(-rowname)

zeale_rare_mat20 <- zeale_rare_mat %>%
  filter(rowname %notin% unique(c(lowreadsampsC20, lowreadsampsZ20))) %>%
  select(-rowname)

# get lowest number of reads
minreads20 <- min(c(coleop_rare_mat20$sample.sum, zeale_rare_mat20$sample.sum))

rarefy_coleop20 <-rarefy(select(coleop_rare_mat20,-sample.sum), 
                         minreads20, se = T)
rarefy_zeale20 <- rarefy(select(zeale_rare_mat20,-sample.sum), 
                         minreads20, se = T)

rarefy_coleop_df20 <- rarefy_coleop20 %>%
  t() %>%
  as.data.frame() %>%
  bind_cols(colnames(rarefy_coleop20),.) %>%
  rename(sample = 1) %>%
  mutate(primer = "16S")

rarefy_zeale_df20 <- rarefy_zeale20 %>%
  t() %>%
  as.data.frame() %>%
  bind_cols(colnames(rarefy_zeale20),.) %>%
  rename(sample = 1) %>%
  mutate(primer = "COI")

rarefy_df20 <- rarefy_coleop_df20 %>%
  bind_rows(rarefy_zeale_df20) %>%
  mutate(method = stringr::str_sub(sample, -1,-1),
         tree = stringr::str_sub(sample, -2,-2)) %>%
  mutate(method = case_when(method == "R" ~ "Bark",
                            TRUE ~ "Understory"),
         tree = case_when(tree == "D" ~ "Deciduous",
                          TRUE ~ "Coniferous"))

# now create a rarified version of the read matrixes for diversity metrics

coleop_rare_mat_rare20 <- rrarefy(select(coleop_rare_mat20,-sample.sum), 
                                  minreads20)

zeale_rare_mat_rare20 <- rrarefy(select(zeale_rare_mat20,-sample.sum), 
                                 minreads20)

# Calculate diversity metrics
coleop_rare_shannon20 <- diversity(coleop_rare_mat_rare20, 
                               index = "shannon", 
                               MARGIN = 1, base = exp(1)) %>%
  as.data.frame() %>%
  bind_cols(rownames(coleop_rare_mat_rare20),.) %>%
  rename(sample = 1, shannon = 2) %>%
  mutate(primer = "16S")

zeale_rare_shannon20 <- diversity(zeale_rare_mat_rare20, 
                               index = "shannon", 
                               MARGIN = 1, base = exp(1)) %>%
  as.data.frame() %>%
  bind_cols(rownames(coleop_rare_mat_rare20),.) %>%
  rename(sample = 1, shannon = 2) %>%
  mutate(primer = "COI")

shannon_df20 <- coleop_rare_shannon20 %>%
  bind_rows(zeale_rare_shannon20) %>%
  mutate(method = stringr::str_sub(sample, -1,-1),
         tree = stringr::str_sub(sample, -2,-2)) %>%
  mutate(method = case_when(method == "R" ~ "Bark",
                            TRUE ~ "Understory"),
         tree = case_when(tree == "D" ~ "Deciduous",
                          TRUE ~ "Coniferous"))

# Calculate Simpsons index on rarefied data
coleop_rare_simpson20 <- diversity(coleop_rare_mat_rare20, 
                               index = "simpson") %>%
  as.data.frame() %>%
  bind_cols(rownames(coleop_rare_mat_rare20),.) %>%
  rename(sample = 1, simpson = 2) %>%
  mutate(primer = "16S")

zeale_rare_simpson20 <- diversity(zeale_rare_mat_rare20, 
                               index = "simpson") %>%
  as.data.frame() %>%
  bind_cols(rownames(zeale_rare_mat_rare20),.) %>%
  rename(sample = 1, simpson = 2) %>%
  mutate(primer = "COI")

simpson_df20 <- coleop_rare_simpson20 %>%
  bind_rows(zeale_rare_simpson20) %>%
  mutate(method = stringr::str_sub(sample, -1,-1),
         tree = stringr::str_sub(sample, -2,-2)) %>%
  mutate(method = case_when(method == "R" ~ "Bark",
                            TRUE ~ "Understory"),
         tree = case_when(tree == "D" ~ "Deciduous",
                          TRUE ~ "Coniferous"))

# set font size for all plot annotations below
annfont <- 2
ytextsize = 10
xtextsize = 11

# Richness model

# run the model by un-commenting this part, or read in the pre-run model below
library(brms) # brms_2.19.0
# (Smod <- brm(log(S) ~ tree + method + primer, data = rarefy_df20, cores = 4,
#              iter = 4000, warmup = 2000))
# saveRDS(Smod, "output/Smod_swarm1_allMOTUs.rds")
(Smod <- readRDS("output/Smod_swarm1_allMOTUs.rds"))
# *** = 99% don't overlap; ** = 95% don't overlap; * = 90% don't overlap
#                  Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept            3.80      0.11     3.58     4.02 1.00    10684     5991
# treeDeciduous        0.19      0.11    -0.03     0.41 1.00    10607     6326*
# methodUnderstory     0.53      0.12     0.30     0.77 1.00    10471     5717***
# primerCOI           -0.80      0.11    -1.03    -0.57 1.00     9444     6128***

# check credible intervals for parameters
library(tidybayes) # tidybayes_3.0.4
quantile(spread_draws(Smod, b_treeDeciduous)$b_treeDeciduous, c(0.05, 0.95)) 
# 0.003040278 0.375500723
quantile(spread_draws(Smod, b_methodUnderstory)$b_methodUnderstory, c(0.005, 0.995)) 
# 0.2166562 0.8642563
quantile(spread_draws(Smod, b_primerCOI)$b_primerCOI, c(0.005, 0.995)) 
# -1.1010535 -0.4997238

# format model predictions for plotting
Smod_df <- conditional_effects(Smod, prob = 0.95) %>%
  do.call(rbind, .) %>%
  mutate(var = c("Tree type", "Tree type", "Substrate", "Substrate", "Primer set", "Primer set"),
         effect1__ = case_when(effect1__ == "Coniferous" ~ "Conifer",
                              TRUE ~ effect1__)) %>%
  rename(q97.5 = upper__,
         q2.5 = lower__) %>%
  bind_cols(select(do.call(rbind, conditional_effects(Smod, prob = 0.80)), q10 = lower__, q90 = upper__))

# create annotation datafame for plotting
Smod_ann <- data.frame(var = c("Tree type", "Substrate", "Primer set"),
                       x = 1.5,
                       y = 2.25,
                       label = c("*", "***", "***"))

# make the predicted MOTU richness plot
(Splot <- 
Smod_df %>%
  ggplot() +
  geom_point(aes(x=effect1__, y = exp(estimate__)),
             size = 3) +
  geom_errorbar(aes(x = effect1__, ymin = exp(q2.5), ymax = exp(q97.5)), width = 0) +
  geom_errorbar(aes(x = effect1__, ymin = exp(q10), ymax = exp(q90)), width = 0, linewidth = 1) +
  geom_text(aes(x = x, y = y, label = label), data = Smod_ann, size = 6) +
  facet_wrap(~var, scales = "free_x") +
  labs(y = "Rarefied MOTU richness", x = "") +
  theme_bw()
)

# Evenness model

# run the model by un-commenting this part, or read in the pre-run model below
# (Simpmod <- brm(simpson ~ tree + method + primer, data = simpson_df20, cores = 4,
#                  iter = 4000, warmup = 2000))
# saveRDS(Simpmod, "output/Simpmod_swarm1_allMOTUs.rds")
(Simpmod <- readRDS("output/Simpmod_swarm1_allMOTUs.rds"))
#                  Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept            0.51      0.07     0.38     0.65 1.00     9659     6105
# treeDeciduous        0.09      0.07    -0.04     0.23 1.00     9562     5408
# methodUnderstory     0.12      0.07    -0.02     0.27 1.00    10467     5913
# primerCOI           -0.14      0.07    -0.28    -0.00 1.00    10333     5709*

# check credible intervals for parameters
get_variables(Simpmod)
quantile(spread_draws(Simpmod, b_treeDeciduous)$b_treeDeciduous, c(0.05, 0.95)) 
# -0.0238576  0.2107642 
quantile(spread_draws(Simpmod, b_methodUnderstory)$b_methodUnderstory, c(0.05, 0.95))
# -0.002474706  0.242093545
quantile(spread_draws(Simpmod, b_primerCOI)$b_primerCOI, c(0.05, 0.95))
# -0.25922331 -0.02676371

# format model predictions for plotting
Simpmod_df <- conditional_effects(Simpmod, prob = 0.95) %>%
  do.call(rbind, .) %>%
  mutate(var = c("Tree type", "Tree type", "Substrate", "Substrate", "Primer set", "Primer set"),
         effect1__ = case_when(effect1__ == "Coniferous" ~ "Conifer",
                              TRUE ~ effect1__)) %>%
  rename(q97.5 = upper__,
         q2.5 = lower__) %>%
  bind_cols(select(do.call(rbind, conditional_effects(Simpmod, prob = 0.80)), q10 = lower__, q90 = upper__))

# create annotation datafame for plotting
Simpmod_ann <- data.frame(var = c("Tree type", "Substrate", "Primer set"),
                       x = 1.5,
                       y = .15,
                       label = c("", "", "*"))

# make the predicted MOTU evenness plot
(Simpplot <-
Simpmod_df %>%
  ggplot() +
  geom_point(aes(x=effect1__, y = estimate__),
             size = 3) +
  geom_errorbar(aes(x = effect1__, ymin = q2.5, ymax = q97.5), width = 0) +
  geom_errorbar(aes(x = effect1__, ymin = q10, ymax = q90), width = 0, size = 1) +
  geom_text(aes(x = x, y = y, label = label), data = Simpmod_ann, size = 6) +
  facet_wrap(~var, scales = "free_x") +
  labs(y = "Rarefied MOTU evenness (Simpson)", x = "") +
  theme_bw()
)
  
# Diversity model

# run the model by un-commenting this part, or read in the pre-run model below
(Shanmod <- brm(shannon ~ tree + method + primer, data = shannon_df20, cores = 4,
                iter = 4000, warmup = 2000))
# saveRDS(Shanmod, "output/Shanmod_swarm1_allMOTUs.rds")
Shanmod <- readRDS("output/Shanmod_swarm1_allMOTUs.rds")
#                  Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept            1.25      0.20     0.86     1.64 1.00     9310     6204
# treeDeciduous        0.33      0.21    -0.08     0.74 1.00    10211     5934
# methodUnderstory     0.64      0.22     0.22     1.07 1.00    10291     6194***
# primerCOI           -0.43      0.21    -0.83    -0.01 1.00     9127     5494**

# check credible intervals for parameters
get_variables(Shanmod)
quantile(spread_draws(Shanmod, b_treeDeciduous)$b_treeDeciduous, c(0.05, 0.95)) 
# -0.008402714  0.676219516
quantile(spread_draws(Shanmod, b_methodUnderstory)$b_methodUnderstory, c(0.995, 0.005))
# 1.21770328 0.08880448
quantile(spread_draws(Shanmod, b_primerCOI)$b_primerCOI, c(0.05, 0.95))
# -0.77219520 -0.08139197 
quantile(spread_draws(Shanmod, b_primerCOI)$b_primerCOI, c(0.005, 0.995))
# -0.9992863  0.1204389

# format model predictions for plotting
Shanmod_df <- conditional_effects(Shanmod, prob = 0.95) %>%
  do.call(rbind, .) %>%
  mutate(var = c("Tree type", "Tree type", "Substrate", "Substrate", "Primer set", "Primer set"),
         effect1__ = case_when(effect1__ == "Coniferous" ~ "Conifer",
                              TRUE ~ effect1__)) %>%
  rename(q97.5 = upper__,
         q2.5 = lower__) %>%
  bind_cols(select(do.call(rbind, conditional_effects(Shanmod, prob = 0.80)), q10 = lower__, q90 = upper__))

# create annotation datafame for plotting
Shanmod_ann <- data.frame(var = c("Tree type", "Substrate", "Primer set"),
                       x = 1.5,
                       y = .1,
                       label = c("", "***", "**"))

# make the predicted MOTU shannon diversity plot
(Shanplot <-
Shanmod_df %>%
  ggplot() +
  geom_point(aes(x=effect1__, y = estimate__),
             size = 3) +
  geom_errorbar(aes(x = effect1__, ymin = q2.5, ymax = q97.5), width = 0) +
  geom_errorbar(aes(x = effect1__, ymin = q10, ymax = q90), width = 0, linewidth = 1) +
  geom_text(aes(x = x, y = y, label = label), data = Shanmod_ann, size = 6) +
  facet_wrap(~var, scales = "free_x") +
  labs(y = "Rarefied MOTU diversity (Shannon)", x = "") +
  theme_bw()
)

library(patchwork)

# plot of marginal effects
gg = gridExtra::arrangeGrob(Shanplot, Splot, Simpplot)

plot(gg)

# ggsave(plot = gg, filename = "figures/Figure3.png",
#        height = 9, width = 4, dpi = 400)
```
# Create Figure 4 - part 1 (left panels)
Create Figure 4 in Allen et al. (2023); just the left panels. PERMANOVA analyses and NMDS plots to evaluate community composition by primer set, substrate, and tree type (conifer vs. deciduous).
```{r}
# format data matrices for species, genus, and family presence/absence

mds.data.sp <- rbind(sprayZ_sp_mat,
                  rollZ_sp_mat,
                  sprayC_sp_mat,
                  rollC_sp_mat)

mds.data.gen <- rbind(sprayZ_gen_mat,
                  rollZ_gen_mat,
                  sprayC_gen_mat,
                  rollC_gen_mat)

mds.data.fam <- rbind(sprayZ_fam_mat,
                  rollZ_fam_mat,
                  sprayC_fam_mat,
                  rollC_fam_mat)

# check for blank rows
sort(apply(mds.data.sp, 1, max)) # sample.9DR_zeale has no species
sort(apply(mds.data.gen, 1, max))
sort(apply(mds.data.fam, 1, max))

# remove blank rows (no blanks)
  # these result from samples where no MOTUs were ID'd to this taxonomic level

mds.data.sp.nb <- mds.data.sp[-which(rownames(mds.data.sp) %in% 
                                     c("sample.9DR_zeale")),]
mds.data.gen.nb <- mds.data.gen
mds.data.fam.nb <- mds.data.fam

### Species level

# format covariate data - species level
env.data <- data.frame(
  tree = str_sub(gsub(gsub(rownames(mds.data.sp.nb), 
                           pattern="_zeale", replacement = ""),
                      pattern="_coleop", replacement = ""), 
                                      -2, -2),
  method = str_sub(gsub(gsub(rownames(mds.data.sp.nb), 
                           pattern="_zeale", replacement = ""),
                      pattern="_coleop", replacement = ""), 
                                        -1, -1)) %>%
  mutate(tree = case_when(tree == "C" ~ "Conifer",
                          tree == "D" ~ "Decid."),
         method = case_when(method == "R" ~ "Bark",
                            method == "S" ~ "Leaves"),
  primer = case_when(grepl(rownames(mds.data.sp.nb), pattern = "zeal") ~ "COI",
                     TRUE ~ "16S"))

# package vegan NMDS
# to run the NMDS, un-comment the code; otherwise load the pre-run model below
set.seed(110)
# arthNMDS.sp = metaMDS(mds.data.sp.nb,
#                      k=2, distance = 'jaccard', trymax = 10000)
# saveRDS(arthNMDS.sp, "output/arthNMDS.sp.swarm1_allMOTUs.rds")
arthNMDS.sp <- readRDS("output/arthNMDS.sp.swarm1_allMOTUs.rds")
arthNMDS.sp # stress: 0.1209

plot1 <- ordiplot(arthNMDS.sp, choices = c(1, 2))

sites.long1 <-
  BiodiversityR::sites.long(plot1, env.data = env.data) %>%
  mutate(axis1 = as.numeric(axis1),
         axis2 = as.numeric(axis2),
         Method = paste0(method, " - ", tree, " (", primer, ")"))
head(sites.long1)
str(sites.long1)

(
  plot.sp <-
    ggplot(sites.long1) +
    geom_vline(
      xintercept = c(0),
      color = "grey70",
      linetype = 2
    ) +
    geom_hline(
      yintercept = c(0),
      color = "grey70",
      linetype = 2
    ) +
    scale_x_continuous(sec.axis = dup_axis(labels = NULL, name = NULL)) +
    scale_y_continuous(sec.axis = dup_axis(labels = NULL, name = NULL)) +
    scale_color_manual(values = c("steelblue", "steelblue", "steelblue", "steelblue", 
                                  "firebrick","firebrick", "firebrick", "firebrick")) +
    scale_fill_manual(values = c("white", "steelblue", "white", "steelblue", 
                                  "white", "firebrick", "white", "firebrick")) +
    scale_shape_manual(values = c(24, 24, 21, 21, 24, 24, 21, 21)) +
    geom_point(
      aes(
        x = axis1,
        y = axis2,
        shape = Method,
        color = Method,
        fill = Method,
      ),
      size = 3,
      alpha = 0.85
    ) +
    annotate("text", x = -1.75, y = -1.1, label = "Stress = 0.121", hjust = 0, size = 3) +
    # coord_fixed(ratio = 1) +
    labs(
      y = "NMDS2",
      x = "NMDS1",
      color = "Substrate",
      shape = "Tree type"
    ) +
    theme_bw() +
    theme(text = element_text(size = 12)) +
    ggtitle("Species") +
   guides(shape = "none", color = "none", fill = "none")
)


# Test for differences among Method and Tree type
# PERMANOVA
set.seed(12555)
perm_arth <- adonis2(mds.data.sp.nb ~ primer + method + tree, data = env.data, 
       permutations = 9999, method = "jaccard",
       strata = NULL, contr.unordered = "contr.sum",
       contr.ordered = "contr.poly")

perm_arth
#          Df SumOfSqs      R2      F Pr(>F)    
# primer    1   3.5329 0.11218 9.1783 0.0001 ***
# method    1   1.2711 0.04036 3.3022 0.0001 ***
# tree      1   0.5159 0.01638 1.3402 0.0615 .  
# Residual 68  26.1746 0.83109                  
# Total    71  31.4945 1.00000          

### Genus level ###

# format covariate data - genus level
env.data.gen <- data.frame(
  tree = str_sub(gsub(gsub(rownames(mds.data.gen.nb), 
                           pattern="_zeale", replacement = ""),
                      pattern="_coleop", replacement = ""), 
                                      -2, -2),
  method = str_sub(gsub(gsub(rownames(mds.data.gen.nb), 
                           pattern="_zeale", replacement = ""),
                      pattern="_coleop", replacement = ""), 
                                        -1, -1)) %>%
  mutate(tree = case_when(tree == "C" ~ "Conifer",
                          tree == "D" ~ "Decid."),
         method = case_when(method == "R" ~ "Bark",
                            method == "S" ~ "Leaves"),
  primer = case_when(grepl(rownames(mds.data.gen.nb), pattern = "zeal") ~ "COI",
                     TRUE ~ "16S"))

# package vegan NMDS
# to run the NMDS, un-comment the code; otherwise load the pre-run model below
set.seed(110)
# arthNMDS.gen = metaMDS(mds.data.gen.nb,
#                      k=2, distance = 'jaccard', trymax = 10000)
# saveRDS(arthNMDS.gen, "output/arthNMDS.gen.swarm1_allMOTUs.rds")
arthNMDS.gen <- readRDS("output/arthNMDS.gen.swarm1_allMOTUs.rds")
arthNMDS.gen # stress = 0.1604

plot1.gen <- ordiplot(arthNMDS.gen, choices = c(1, 2))

sites.long1.gen <-
  BiodiversityR::sites.long(plot1.gen, env.data = env.data.gen) %>%
  mutate(axis1 = as.numeric(axis1),
         axis2 = as.numeric(axis2),
         Method = paste0(method, " - ", tree, " (", primer, ")"))
head(sites.long1.gen)
str(sites.long1.gen)

(
  plot.gen <-
    ggplot(sites.long1.gen) +
    geom_vline(
      xintercept = c(0),
      color = "grey70",
      linetype = 2
    ) +
    geom_hline(
      yintercept = c(0),
      color = "grey70",
      linetype = 2
    ) +
    scale_x_continuous(sec.axis = dup_axis(labels = NULL, name = NULL)) +
    scale_y_continuous(sec.axis = dup_axis(labels = NULL, name = NULL)) +
    scale_color_manual(values = c("steelblue", "steelblue", "steelblue", "steelblue", 
                                  "firebrick","firebrick", "firebrick", "firebrick")) +
    scale_fill_manual(values = c("white", "steelblue", "white", "steelblue", 
                                  "white", "firebrick", "white", "firebrick")) +
    scale_shape_manual(values = c(24, 24, 21, 21, 24, 24, 21, 21)) +
    geom_point(
      aes(
        x = axis1,
        y = axis2,
        shape = Method,
        color = Method,
        fill = Method
      ),
      size = 3,
      alpha = 0.85
    ) +
    annotate("text", x = -2.5, y = -0.85, label = "Stress = 0.160", hjust = 0, size = 3) +
    # coord_fixed(ratio = 1) +
    labs(
      y = "NMDS2",
      x = "NMDS1",
      color = "Substrate",
      shape = "Tree type"
    ) +
    theme_bw() +
    theme(text = element_text(size = 12)) +
    ggtitle("Genus") +
    guides(shape = "none", color = "none", fill = "none")
)

# Test for differences among Method and Tree type (genus level)
# PERMANOVA
set.seed(12555)
perm_arth.gen <- adonis2(mds.data.gen.nb ~ primer + method + tree, data = env.data.gen, 
       permutations = 9999, method = "jaccard",
       strata = NULL, contr.unordered = "contr.sum",
       contr.ordered = "contr.poly")

perm_arth.gen
#          Df SumOfSqs      R2      F Pr(>F)    
# primer    1   2.7781 0.09070 7.4135 0.0001 ***
# method    1   1.4912 0.04868 3.9793 0.0001 ***
# tree      1   0.5048 0.01648 1.3470 0.0434 *  
# Residual 69  25.8564 0.84414                  
# Total    72  30.6304 1.00000 

### Family level ###

# format covariate data - family level
env.data.fam <- data.frame(
  tree = str_sub(gsub(gsub(rownames(mds.data.fam.nb), 
                           pattern="_zeale", replacement = ""),
                      pattern="_coleop", replacement = ""), 
                                      -2, -2),
  method = str_sub(gsub(gsub(rownames(mds.data.fam.nb), 
                           pattern="_zeale", replacement = ""),
                      pattern="_coleop", replacement = ""), 
                                        -1, -1)) %>%
  mutate(tree = case_when(tree == "C" ~ "Conifer",
                          tree == "D" ~ "Decid."),
         method = case_when(method == "R" ~ "Bark",
                            method == "S" ~ "Leaves"),
  primer = case_when(grepl(rownames(mds.data.fam.nb), pattern = "zeal") ~ "COI",
                     TRUE ~ "16S"))

# package vegan NMDS
# to run the NMDS, un-comment the code; otherwise load the pre-run model below
set.seed(110)
# arthNMDS.fam = metaMDS(mds.data.fam.nb,
#                      k=2, distance = 'jaccard', trymax = 10000)
# saveRDS(arthNMDS.fam, "output/arthNMDS.fam.swarm1_allMOTUs.rds")
arthNMDS.fam <- readRDS("output/arthNMDS.fam.swarm1_allMOTUs.rds")
arthNMDS.fam # stress = 0.2043

plot1.fam <- ordiplot(arthNMDS.fam, choices = c(1, 2))

sites.long1.fam <-
  BiodiversityR::sites.long(plot1.fam, env.data = env.data.fam) %>%
  mutate(axis1 = as.numeric(axis1),
         axis2 = as.numeric(axis2),
         Method = paste0(method, " - ", tree, " (", primer, ")"))
head(sites.long1.fam)
str(sites.long1.fam)

(
  plot.fam <-
    ggplot(sites.long1.fam) +
    geom_vline(
      xintercept = c(0),
      color = "grey70",
      linetype = 2
    ) +
    geom_hline(
      yintercept = c(0),
      color = "grey70",
      linetype = 2
    ) +
    scale_x_continuous(sec.axis = dup_axis(labels = NULL, name = NULL)) +
    scale_y_continuous(sec.axis = dup_axis(labels = NULL, name = NULL)) +
    scale_color_manual(values = c("steelblue", "steelblue", "steelblue", "steelblue", 
                                  "firebrick","firebrick", "firebrick", "firebrick")) +
    scale_fill_manual(values = c("white", "steelblue", "white", "steelblue", 
                                  "white", "firebrick", "white", "firebrick")) +
    scale_shape_manual(values = c(24, 24, 21, 21, 24, 24, 21, 21)) +
    geom_point(
      aes(
        x = axis1,
        y = axis2,
        shape = Method,
        color = Method,
        fill = Method
      ),
      size = 3,
      alpha = 0.85
    ) +
    annotate("text", x = -1.8, y = -0.7, label = "Stress = 0.204", hjust = 0, size = 3) +
    # coord_fixed(ratio = 1) +
    labs(
      y = "NMDS2",
      x = "NMDS1",
      color = "Substrate",
      shape = "Tree type"
    ) +
    theme_bw() +
    theme(text = element_text(size = 12)) +
    ggtitle("Family") +
    guides(shape = "none", color = "none", fill = "none")
)

# Test for differences among Method and Tree type
# PERMANOVA
set.seed(12555)
perm_arth.fam <- adonis2(mds.data.fam.nb ~ primer + method + tree, data = env.data.fam, 
       permutations = 9999, method = "jaccard",
       strata = NULL, contr.unordered = "contr.sum",
       contr.ordered = "contr.poly")

perm_arth.fam
#          Df SumOfSqs      R2       F Pr(>F)    
# primer    1   3.2696 0.13060 11.4864 0.0001 ***
# method    1   1.6391 0.06547  5.7582 0.0001 ***
# tree      1   0.4852 0.01938  1.7046 0.0164 *  
# Residual 69  19.6407 0.78454                   
# Total    72  25.0345 1.00000   

# plot will be saved in the next step ('create Figure 4 - part 2')
```
# Create Figure 4 - part 2 (right panels)
Create Figure 4 in Allen et al. (2023); just the right panels. Count up taxa by primer set and substrate and make Venn diagrams.
```{r}

### family

# subdivide 'art' into columns by primer / substrate / tree type for counting
Z_spray <- art_fam_reads %>%
  select(family, contains(c("CS_zeale", "DS_zeale"))) %>%
   mutate(sum = rowSums(across(where(is.numeric)))) %>%
  filter(sum > 0) %>%
  select(family) %>%
  distinct() 

Z_roll <- art_fam_reads %>%
  select(family, contains(c("CR_zeale", "DR_zeale"))) %>%
   mutate(sum = rowSums(across(where(is.numeric)))) %>%
  filter(sum > 0) %>%
  select(family) %>%
  distinct() 

C_spray <- art_fam_reads %>%
  select(family, contains(c("CS_coleop", "DS_coleop"))) %>%
   mutate(sum = rowSums(across(where(is.numeric)))) %>%
  filter(sum > 0) %>%
  select(family) %>%
  distinct() 

C_roll <- art_fam_reads %>%
  select(family, contains(c("CR_coleop", "DR_coleop"))) %>%
   mutate(sum = rowSums(across(where(is.numeric)))) %>%
  filter(sum > 0) %>%
  select(family) %>%
  distinct() 

# get unique families within each method (change names to get different combos)
# zeale has 117 families (54 of which are not found in coleop)
# coleop has 127 families (64 of which are not found in zeale)
(Z_uni <- Z_spray %>%
    bind_rows(Z_roll) %>%
    distinct() %>%
  filter(family %notin% unique(c(
                           C_spray$family,
                           C_roll$family)))) 

(C_uni <- C_spray %>%
    bind_rows(C_roll) %>%
    distinct() %>%
  filter(family %notin% unique(c(
                           Z_spray$family,
                           Z_roll$family)))) 

# Venn diagram of the 8 primer/substrates
# code from: https://www.datanovia.com/en/blog/venn-diagram-with-r-or-rstudio-a-million-ways/
# installed using the following code:
# devtools::install_github("yanlinlin82/ggvenn")

venn_methods_fam <- list(
  'US\n(COI)' = Z_spray$family,
  'US (16S)' = C_spray$family,
  'Bark (COI)' = Z_roll$family,
  'Bark\n(16S)' = C_roll$family
)

library(ggVennDiagram) # ggVennDiagram_1.2.3
(vennfam <- ggVennDiagram(venn_methods_fam, label_alpha = 0, edge_lty = 1,
              label_color = "black", edge_size = 0.5, set_size = 3, label_size = 3) +
  scale_fill_gradient(low="pink",high = "firebrick") +
  scale_color_manual(values = c(rep("darkgray", 4))) +
  labs(fill = "")
)

# pairwise overlap (manually counted from Venn diagram)
# leaves-bark: 96 overlapping of 181 total (53% overlap)
unique(c(Z_spray$family, C_spray$family))[unique(c(Z_spray$family, C_spray$family)) %in% unique(c(Z_roll$family, C_roll$family))]
# COI-16S: 62 overlapping of 181 total (34% overlap)
unique(c(Z_spray$family, Z_roll$family))[unique(c(Z_spray$family, Z_roll$family)) %in% unique(c(C_spray$family, C_roll$family))]
# conif-decid: xx overlapping of 181 total (xx% overlap)

### genus

# subdivide 'art' into columns by primer / substrate / tree type for counting
Z_spray.gen <- art_gen_reads %>%
  select(genus, contains(c("CS_zeale", "DS_zeale"))) %>%
   mutate(sum = rowSums(across(where(is.numeric)))) %>%
  filter(sum > 0) %>%
  select(genus) %>%
  distinct() # 169 for zeale/spray

Z_roll.gen <- art_gen_reads %>%
  select(genus, contains(c("CR_zeale", "DR_zeale"))) %>%
   mutate(sum = rowSums(across(where(is.numeric)))) %>%
  filter(sum > 0) %>%
  select(genus) %>%
  distinct() # 118 for zeale/roll

C_spray.gen <- art_gen_reads %>%
  select(genus, contains(c("CS_coleop", "DS_coleop"))) %>%
   mutate(sum = rowSums(across(where(is.numeric)))) %>%
  filter(sum > 0) %>%
  select(genus) %>%
  distinct() # 145 for coleop/spray

C_roll.gen <- art_gen_reads %>%
  select(genus, contains(c("CR_coleop", "DR_coleop"))) %>%
   mutate(sum = rowSums(across(where(is.numeric)))) %>%
  filter(sum > 0) %>%
  select(genus) %>%
  distinct() # 88 for coleop/roll

# get unique genera within each method (change names to get different combos)
# zeale has 227 genera (180 of which are not found in coleop)
# coleop has 173 genera (126 of which are not found in zeale)
(C_uni.gen <- C_spray.gen %>%
    bind_rows(C_roll.gen) %>%
    distinct() %>%
  filter(genus %notin% unique(c(
                           Z_spray.gen$genus, 
                           Z_roll.gen$genus)))) 

(Z_uni.gen <- Z_spray.gen %>%
    bind_rows(Z_roll.gen) %>%
    distinct() %>%
  filter(genus %notin% unique(c(
                           C_spray.gen$genus, 
                           C_roll.gen$genus)))) 

# Venn diagram of the 8 primer/substrates

venn_methods_gen <- list(
  'US\n(COI)' = Z_spray.gen$genus,
  'US (16S)' = C_spray.gen$genus,
  'Bark (COI)' = Z_roll.gen$genus,
  'Bark\n(16S)' = C_roll.gen$genus
)

(venngen <- ggVennDiagram(venn_methods_gen, label_alpha = 0, edge_lty = 1,
              label_color = "black", edge_size = 0.5, set_size = 3, label_size = 3) +
  scale_fill_gradient(low="pink",high = "firebrick") +
  scale_color_manual(values = c(rep("darkgray", 4))) +
  labs(fill = "")
)

# pairwise overlap (manually counted from Venn diagram)
# leaves-bark: 109 overlapping of 353 total (31% overlap)
unique(c(Z_spray.gen$genus, C_spray.gen$genus))[unique(c(Z_spray.gen$genus, C_spray.gen$genus)) %in% unique(c(Z_roll.gen$genus, C_roll.gen$genus))]
# COI-16S: 47 overlapping of 353 total (13% overlap)
unique(c(Z_spray.gen$genus, Z_roll.gen$genus))[unique(c(Z_spray.gen$genus, Z_roll.gen$genus)) %in% unique(c(C_roll.gen$genus, C_spray.gen$genus))]

### species

# subdivide 'art' into columns by primer / substrate / tree type for counting
Z_spray.sp <- art_sp_reads %>%
  select(species, contains(c("CS_zeale", "DS_zeale"))) %>%
   mutate(sum = rowSums(across(where(is.numeric)))) %>%
  filter(sum > 0) %>%
  select(species) %>%
  distinct() # 149 for zeale/spray

Z_roll.sp <- art_sp_reads %>%
  select(species, contains(c("CR_zeale", "DR_zeale"))) %>%
   mutate(sum = rowSums(across(where(is.numeric)))) %>%
  filter(sum > 0) %>%
  select(species) %>%
  distinct() # 87 for zeale/roll

C_spray.sp <- art_sp_reads %>%
  select(species, contains(c("CS_coleop", "DS_coleop"))) %>%
   mutate(sum = rowSums(across(where(is.numeric)))) %>%
  filter(sum > 0) %>%
  select(species) %>%
  distinct() # 98 for coleop/spray

C_roll.sp <- art_sp_reads %>%
  select(species, contains(c("CR_coleop", "DR_coleop"))) %>%
   mutate(sum = rowSums(across(where(is.numeric)))) %>%
  filter(sum > 0) %>%
  select(species) %>%
  distinct() # 54 for coleop/roll

# get unique genera within each method (change names to get different combos)
# zeale has 197 sp (177 of which are not found in coleop)
# coleop has 115 sp (95 of which are not found in zeale)
# so they only share 20 species in common, zeale has 176 unique sp, coleop has 90 unique sp
(C_uni.sp <- C_spray.sp %>%
    bind_rows(C_roll.sp) %>%
    distinct() %>%
  filter(species %notin% unique(c(
                           Z_spray.sp$species, 
                           Z_roll.sp$species)))) 

(Z_uni.sp <- Z_spray.sp %>%
    bind_rows(Z_roll.sp) %>%
    distinct() %>%
  filter(species %notin% unique(c(
                           C_spray.sp$species, 
                           C_roll.sp$species)))) 

# Venn diagram of the 8 primer/substrates

venn_methods_sp <- list(
  'US\n(COI)' = Z_spray.sp$species,
  'US (16S)' = C_spray.sp$species,
  'Bark (COI)' = Z_roll.sp$species,
  'Bark\n(16S)' = C_roll.sp$species
)

(vennsp <- ggVennDiagram(venn_methods_sp, label_alpha = 0, edge_lty = 1,
              label_color = "black", edge_size = 0.5, set_size = 3, label_size = 3) +
  scale_fill_gradient(low="pink",high = "firebrick") +
  scale_color_manual(values = c(rep("darkgray", 4))) +
  labs(fill = "")
)

# pairwise overlap (manually counted from Venn diagram)
# leaves-bark: 71 overlapping of 292 total (24% overlap)
unique(c(Z_spray.sp$species, C_spray.sp$species))[unique(c(Z_spray.sp$species, C_spray.sp$species)) %in% unique(c(Z_roll.sp$species, C_roll.sp$species))]
# COI-16S: 20 overlapping of 292 total (7% overlap)
unique(c(Z_spray.sp$species, Z_roll.sp$species))[unique(c(Z_spray.sp$species, Z_roll.sp$species)) %in% unique(c(C_spray.sp$species, C_roll.sp$species))]

# combine all plots from previous section (left panels) and this section (right panels) into Fig. 4
library(patchwork)

plot.sp + plot_spacer() + vennsp + plot.gen + plot_spacer() + venngen + plot.fam + plot_spacer() + vennfam + 
  plot_layout(ncol = 3, widths = c(1,.1,1,1,.1,1,1,.1,1))

# ggsave(filename = "figures/Figure4.png",
#        height = 10, width = 7, dpi = 400)
```
# Create Figure 5 - part 1 of 4
Create Figure 5 from Allen et al. (2023); part 1 of 4. Run multinomial models for diet/feeding guild data. Code adapted from these sources:
https://discourse.mc-stan.org/t/example-with-family-multinomial/8707
https://discourse.mc-stan.org/t/interpret-parameters-from-multinomial-categorical-logit-regression/26374
```{r}
# format diet data for multinomial model
fam_pa <- art_fam_motus[,-1]
fam_pa[is.na(fam_pa)] <- 0
fam_pa <- cbind(art_fam_motus[,1], fam_pa)

feed_freq <- fam_pa %>%
  left_join(select(diet, family, diet.cat2), by = join_by(family)) %>%
  select(-family, -starts_with("sample.sum"), -contains("3CS_zeale"))

feed_freq_sum1 <- aggregate(feed_freq[,1:73], 
                           by = list(feed_freq$diet.cat2), sum)

feed_freq_sumC1 <- feed_freq_sum1 %>%
  select(contains("coleop")) %>%
  t()
colnames(feed_freq_sumC1) <- feed_freq_sum1$Group.1

feed_freq_sumZ1 <- feed_freq_sum1 %>%
  select(contains("zeale")) %>%
  t()
colnames(feed_freq_sumZ1) <- feed_freq_sum1$Group.1

feed_freq_sumC <- feed_freq_sumC1 %>%
  as.data.frame() %>%
  mutate(sample = rownames(feed_freq_sumC1),
         tree = stringr::str_sub(sample, -9,-9),
         method = stringr::str_sub(sample, -8, -8),
         primer = "16S") %>%
  rename(mult = 1, myce = 2, para = 3, phyt = 4, pred = 5, scav = 6) %>%
  select(-sample)

feed_freq_sumZ <- feed_freq_sumZ1 %>%
  as.data.frame() %>%
  mutate(sample = rownames(feed_freq_sumZ1),
         tree = stringr::str_sub(sample, -8,-8),
         method = stringr::str_sub(sample, -7, -7),
         primer = "COI") %>%
  rename(mult = 1, myce = 2, para = 3, phyt = 4, pred = 5, scav = 6) %>%
  select(-sample)

feed_freq_sum <- feed_freq_sumC %>%
  bind_rows(feed_freq_sumZ) %>%
  rowwise() %>%
  mutate(scav = sum(c_across(c(scav, myce)))) %>%
  ungroup() %>%
  select(-myce)

rm(feed_freq, feed_freq_sum1, feed_freq_sumC1, feed_freq_sumZ1, feed_freq_sumC, feed_freq_sumZ)

# run the multinomial models for diet/feeding guild 
library(brms)

dat <- feed_freq_sum
dat$a1 <- 1
dat$size <- with(dat, mult + para + phyt + pred + scav)
dat$y <- with(dat, cbind(mult, para, phyt, pred, scav))

# # model with primer, method, and tree type
set.seed(102)
fit.prim.method.tree <- brm(bf(y | trials(size) ~ primer + method + tree),
            data = dat, family = multinomial(refcat = "mult"),
            save_pars = save_pars(all = TRUE), cores = 4,
            warmup = 2000, iter = 4000)

# # model with only primer and method
set.seed(102)
fit.prim.method <- brm(bf(y | trials(size) ~ primer + method),
            data = dat, family = multinomial(refcat = "mult"),
            save_pars = save_pars(all = TRUE), cores = 4,
            warmup = 2000, iter = 4000)

# model with only primer and tree
set.seed(102)
fit.prim.tree <- brm(bf(y | trials(size) ~ primer + tree),
            data = dat, family = multinomial(refcat = "mult"),
            save_pars = save_pars(all = TRUE), cores = 4,
            warmup = 2000, iter = 4000)

# model with only primer
set.seed(102)
fit.prim <- brm(bf(y | trials(size) ~ primer),
            data = dat, family = multinomial(refcat = "mult"),
            save_pars = save_pars(all = TRUE), cores = 4,
            warmup = 2000, iter = 4000)

# adding LOOIC criterion to 4 models
fit.prim.method.tree <- add_criterion(fit.prim.method.tree, "loo",
                                 moment_match = T,
                          reloo = TRUE)
fit.prim.method <- add_criterion(fit.prim.method, "loo",
                                 moment_match = T,
                          reloo = TRUE)
fit.prim.tree <- add_criterion(fit.prim.tree, "loo",
                                 moment_match = T,
                          reloo = TRUE)
fit.prim <- add_criterion(fit.prim, "loo",
                                 moment_match = T,
                          reloo = TRUE)

# evaluating models based on LOOIC
brms::loo_compare(fit.prim.method.tree, fit.prim.method, fit.prim.tree, fit.prim,
            criterion = "loo") %>%
  print(simplify = F)

# primer + method is lowest, nearest model is delta elpd = -3.7 (se = 2.5)
```
# Create Figure 5 - part 2 of 4
Create Figure 5 from Allen et al. (2023); part 2 of 4. Run multinomial models for lifemode data. Code adapted from these sources:
https://discourse.mc-stan.org/t/example-with-family-multinomial/8707
https://discourse.mc-stan.org/t/interpret-parameters-from-multinomial-categorical-logit-regression/26374 
```{r}
# format data for multinomial models
fam_pa <- art_fam_motus[,-1]
fam_pa[is.na(fam_pa)] <- 0
fam_pa <- cbind(art_fam_motus[,1], fam_pa)

life_freq <- fam_pa  %>%
  left_join(select(lifemode, family, life.cat), by = join_by(family)) %>%
  select(-family, -starts_with("sample.sum"), -contains("3CS_zeale"))

life_freq_sum1 <- aggregate(life_freq[,1:73], 
                           by = list(life_freq$life.cat), sum)

life_freq_sumC1 <- life_freq_sum1 %>%
  select(contains("coleop")) %>%
  t()
colnames(life_freq_sumC1) <- life_freq_sum1$Group.1

life_freq_sumZ1 <- life_freq_sum1 %>%
  select(contains("zeale")) %>%
  t()
colnames(life_freq_sumZ1) <- life_freq_sum1$Group.1

life_freq_sumC <- life_freq_sumC1 %>%
  as.data.frame() %>%
  mutate(sample = rownames(life_freq_sumC1),
         tree = stringr::str_sub(sample, -9,-9),
         method = stringr::str_sub(sample, -8, -8),
         primer = "16S") %>%
  rename(aqlar = 1, fullaq = 2, fullter = 3, 
         someaq = 4) %>%
  select(-sample)

life_freq_sumZ <- life_freq_sumZ1 %>%
  as.data.frame() %>%
  mutate(sample = rownames(life_freq_sumZ1),
         tree = stringr::str_sub(sample, -8,-8),
         method = stringr::str_sub(sample, -7, -7),
         primer = "COI") %>%
  rename(aqlar = 1, fullaq = 2, fullter = 3, 
         someaq = 4) %>%
  select(-sample)

# format final data file
life_freq_sum <- life_freq_sumC %>%
  bind_rows(life_freq_sumZ) %>%
  mutate(aqlar = fullaq + aqlar) %>%
  select(-fullaq)

# remove intermediate steps
rm(life_freq, life_freq_sum1, life_freq_sumC1, life_freq_sumZ1, life_freq_sumC, life_freq_sumZ)

# run the multinomial models for lifemode
library(brms)

dat <- life_freq_sum
dat$size <- with(dat, aqlar + fullter + someaq) # note: removed fullaq because few members
dat$y <- with(dat, cbind(aqlar, fullter, someaq)) # note: removed fullaq because few members

# model with primer, method, and tree type
set.seed(102)
life.prim.method.tree <- brms::brm(bf(y | trials(size) ~ primer + method + tree), 
            data = dat, family = multinomial(refcat = "aqlar"),
            save_pars = save_pars(all = TRUE), 
                          iter = 4000, warmup = 2000,
                          chains = 4, cores = 4)

# model with primer, method
set.seed(102)
life.prim.method <- brms::brm(bf(y | trials(size) ~ primer + method), 
            data = dat, family = multinomial(refcat = "aqlar"),
            save_pars = save_pars(all = TRUE), 
                          iter = 4000, warmup = 2000,
                          chains = 4, cores = 4)

# model with primer, tree type
set.seed(102)
life.prim.tree <- brms::brm(bf(y | trials(size) ~ primer + tree), 
            data = dat, family = multinomial(refcat = "aqlar"),
            save_pars = save_pars(all = TRUE), 
                          iter = 4000, warmup = 2000,
                          chains = 4, cores = 4)

# model with just primer
set.seed(102)
life.prim <- brms::brm(bf(y | trials(size) ~ primer), 
            data = dat, family = multinomial(refcat = "aqlar"),
            save_pars = save_pars(all = TRUE), 
                          iter = 4000, warmup = 2000,
                          chains = 4, cores = 4)

# adding LOOIC criterion to 4 models
life.prim.method.tree <- brms::add_criterion(life.prim.method.tree, "loo",
                                 moment_match = T,
                          reloo = TRUE)
life.prim.method <- brms::add_criterion(life.prim.method, "loo",
                                 moment_match = T,
                          reloo = TRUE)
life.prim.tree <- brms::add_criterion(life.prim.tree, "loo",
                                 moment_match = T,
                          reloo = TRUE)
life.prim <- brms::add_criterion(life.prim, "loo",
                                 moment_match = T,
                          reloo = TRUE)

# evaluating models based on LOOIC
brms::loo_compare(life.prim.method.tree, life.prim.method, life.prim.tree, life.prim,
            criterion = "loo") %>%
  print(simplify = F)

# primer + method is lowest, delta elpd to next lowest model = -0.9 (se 2.5)
```
# Create Figure 5 - part 3 of 4 
Create Figure 5 from Allen et al. (2023); part 3 of 4. Run multinomial models for size data. Code adapted from these sources:
https://discourse.mc-stan.org/t/example-with-family-multinomial/8707
https://discourse.mc-stan.org/t/interpret-parameters-from-multinomial-categorical-logit-regression/26374
```{r}
# load data for multinomial models
fam_pa <- art_fam_motus[,-1]
fam_pa[is.na(fam_pa)] <- 0
fam_pa <- cbind(art_fam_motus[,1], fam_pa)

size_freq <- fam_pa  %>%
  left_join(select(size, family, size.cat = Body_Size), by = join_by(family)) %>%
  select(-family, -starts_with("sample.sum"), -contains("3CS_zeale"))

size_freq_sum1 <- aggregate(size_freq[,1:73], 
                           by = list(size_freq$size.cat), sum)

size_freq_sumC1 <- size_freq_sum1 %>%
  select(contains("coleop")) %>%
  t()
colnames(size_freq_sumC1) <- size_freq_sum1$Group.1

size_freq_sumZ1 <- size_freq_sum1 %>%
  select(contains("zeale")) %>%
  t()
colnames(size_freq_sumZ1) <- size_freq_sum1$Group.1

size_freq_sumC <- size_freq_sumC1 %>%
  as.data.frame() %>%
  mutate(sample = rownames(size_freq_sumC1),
         tree = stringr::str_sub(sample, -9,-9),
         method = stringr::str_sub(sample, -8, -8),
         primer = "16S") %>%
  rename(vsmal = 1, vbig = 2, big = 3, 
         smal = 4) %>%
  select(-sample)

size_freq_sumZ <- size_freq_sumZ1 %>%
  as.data.frame() %>%
  mutate(sample = rownames(size_freq_sumZ1),
         tree = stringr::str_sub(sample, -8,-8),
         method = stringr::str_sub(sample, -7, -7),
         primer = "COI") %>%
  rename(vsmal = 1, vbig = 2, big = 3, 
         smal = 4) %>%
  select(-sample)

# format final data file
size_freq_sum <- size_freq_sumC %>%
  bind_rows(size_freq_sumZ)

# remove intermediate steps
rm(size_freq, size_freq_sum1, size_freq_sumC1, size_freq_sumZ1, size_freq_sumC, size_freq_sumZ)

# run the multinomial models for lifemode
library(brms)

dat <- size_freq_sum
dat$size <- with(dat, vsmal + vbig + big + smal)
dat$y <- with(dat, cbind(vsmal, vbig, big, smal))

# model with primer, method and tree type
set.seed(102)
size.prim.method.tree <- brms::brm(bf(y | trials(size) ~ primer + method + tree), 
            data = dat, family = multinomial(refcat = "vsmal"),
            save_pars = save_pars(all = TRUE), 
                          iter = 4000, warmup = 2000,
                          chains = 4, cores = 4)

# model with primer, method
set.seed(102)
size.prim.method <- brms::brm(bf(y | trials(size) ~ primer + method), 
            data = dat, family = multinomial(refcat = "vsmal"),
            save_pars = save_pars(all = TRUE), 
                          iter = 4000, warmup = 2000,
                          chains = 4, cores = 4)

# model with primer, tree type
set.seed(102)
size.prim.tree <- brms::brm(bf(y | trials(size) ~ primer + tree), 
            data = dat, family = multinomial(refcat = "vsmal"),
            save_pars = save_pars(all = TRUE), 
                          iter = 4000, warmup = 2000,
                          chains = 4, cores = 4)

# model with just primer
set.seed(102)
size.prim <- brms::brm(bf(y | trials(size) ~ primer), 
            data = dat, family = multinomial(refcat = "vsmal"),
            save_pars = save_pars(all = TRUE), 
                          iter = 4000, warmup = 2000,
                          chains = 4, cores = 4)

# adding LOOIC criterion to 4 models
size.prim.method.tree <- brms::add_criterion(size.prim.method.tree, "loo",
                                 moment_match = T,
                          reloo = TRUE)
size.prim.method <- brms::add_criterion(size.prim.method, "loo",
                                 moment_match = T,
                          reloo = TRUE)
size.prim.tree <- brms::add_criterion(size.prim.tree, "loo",
                                 moment_match = T,
                          reloo = TRUE)
size.prim <- brms::add_criterion(size.prim, "loo",
                                 moment_match = T,
                          reloo = TRUE)

# evaluating models based on LOOIC
brms::loo_compare(size.prim.method.tree, size.prim.method, size.prim.tree, size.prim,
            criterion = "loo") %>%
  print(simplify = F)

# method is lowest; delta elpd to nearest = -2.7 (2.1 se)
```
# Create Figure 5 - part 4 of 4 
Create Figure 5 from Allen et al. (2023); part 4 of 4. Plot multinomial model predictions for diet, lifemode, and size. Note: first, you will need to run at least the "primer + method" models from the 3 sections above. Code adapted from these sources:
https://discourse.mc-stan.org/t/example-with-family-multinomial/8707
https://discourse.mc-stan.org/t/interpret-parameters-from-multinomial-categorical-logit-regression/26374
```{r}
### Feeding guild

### NOTE: first, run at least the "primer + method" models from 3 sections above

# plot settings
dod <- .5
psiz <- 2.5
thin <- 0.3
thik <- 0.75

# plot top model predictions

# format data for diet plot
fit.newdata = expand.grid(method = c("S", "R"),
                          primer = c("COI", "16S"),
                                 size = 1)

fitsies <- fitted(fit.prim.method, 
                  newdata = fit.newdata,probs = c(0.025, 0.1,
                                                  0.9, 0.975))

dim(fitsies )
fitsies # dims are expand grid rows (4), quantiles etc. (6), levels of Y (5)
names(fitsies[1,1,])

fitsies.list <- list()
for(i in 1:dim(fitsies)[3]){
  fitsies.list[[i]] <- fitsies[,,i]
}

fits_df <- do.call(rbind, fitsies.list) %>%
  as.data.frame() %>%
  mutate(method = rep(fit.newdata$method, 5),
         primer = rep(fit.newdata$primer, 5),
         feed = c(rep("Mulitple categories", 4),
                  rep("Parasite/Parasitoid", 4),
                  rep("Phytophagous", 4),
                  rep("Predator", 4),
                  rep("Scavenger", 4))) %>%
  group_by(feed) %>%
  mutate(mean_p = mean(Estimate)) %>%
  ungroup() %>%
  mutate(feed = fct_reorder(feed, mean_p),
         method = case_when(method == "S" ~ "Understory",
                            method == "R" ~ "Bark"))

# create diet plot
(feed_plot2 <-
fits_df %>%
    mutate(methodprimer = paste0(method, " (", primer, ")")) %>%
  ggplot() +
  geom_errorbar(aes(y = feed, xmin = 100*Q2.5, 
                    xmax = 100*Q97.5, color = methodprimer,
                    group = methodprimer),
                position = position_dodge(width = dod),
                width = 0, size = thin) +  
  geom_errorbar(aes(y = feed, xmin = 100*Q10, 
                    xmax = 100*Q90, color = methodprimer,
                    group = methodprimer),
                size = thik,
                position = position_dodge(width = dod),
                width = 0) +
  geom_point(aes(y = feed, x = 100*Estimate, color = methodprimer,
                 fill = methodprimer,
                 shape = methodprimer),
             position = position_dodge(width = dod),
             size = psiz) +
  scale_color_manual(values = c("steelblue", "steelblue", "firebrick", "firebrick")) +
  scale_fill_manual(values = c("steelblue", "white", "firebrick", "white")) +
  scale_shape_manual(values = c(21,21,21,21)) +
  labs(color = "",
       y = "",
       x = "",
       title = "Feeding guild",
       fill = "Sample type",
       shape = "Sample type") +
  theme_bw() +
  theme(text = element_text(size = 13),
        legend.position = c(.8,.2)) +
    guides(color = "none", shape = "none", fill = "none")
 )

### Body size

# plot top model predictions

# format data for size plot
fit.newdata = expand.grid(method = c("S", "R"),
                          primer = c("COI", "16S"),
                                 size = 1)

fitsies <- fitted(size.prim.method, 
                  newdata = fit.newdata,probs = c(0.025, 0.1,
                                                  0.9, 0.975))

dim(fitsies)
fitsies # dims are expand grid rows (4), quantiles etc. (6), levels of Y (4)
names(fitsies[1,1,])

fitsies.list <- list()
for(i in 1:dim(fitsies)[3]){
  fitsies.list[[i]] <- fitsies[,,i]
}

size_fits_df <- do.call(rbind, fitsies.list) %>%
  as.data.frame() %>%
  mutate(method = rep(fit.newdata$method, 4),
         primer = rep(fit.newdata$primer, 4),
         size = c(rep("< 5 mm", 4),
                  rep("> 20 mm", 4),
                  rep("10-20 mm", 4),
                  rep("5-10 mm", 4))) %>%
  group_by(size) %>%
  mutate(mean_p = mean(Estimate)) %>%
  ungroup() %>%
  mutate(size = fct_reorder(size, mean_p),
         method = case_when(method == "S" ~ "Understory",
                            method == "R" ~ "Bark"))

# creaet size plot
(size_plot2 <- 
size_fits_df %>%
    mutate(methodtreeprimer = paste0(method, " (", primer, ")")) %>%
  ggplot() +
  geom_errorbar(aes(y = size, xmin = 100*Q2.5, 
                    xmax = 100*Q97.5, color = methodtreeprimer,
                    group = methodtreeprimer),
                position = position_dodge(width = dod),
                width = 0, size = thin) +  
  geom_errorbar(aes(y = size, xmin = 100*Q10, 
                    xmax = 100*Q90, color = methodtreeprimer,
                    group = methodtreeprimer),
                size = thik,
                position = position_dodge(width = dod),
                width = 0) +
  geom_point(aes(y = size, x = 100*Estimate, color = methodtreeprimer,
                 fill = methodtreeprimer,
                 shape = methodtreeprimer),
             position = position_dodge(width = dod),
             size = psiz) +
  scale_color_manual(values = c("steelblue", "steelblue", "firebrick", "firebrick")) +
  scale_fill_manual(values = c("steelblue", "white", "firebrick", "white")) +
  scale_shape_manual(values = c(21,21,21,21)) +
  labs(color = "",
       y = "",
       x = "",
       title = "Body size",
       fill = "Sample type",
       shape = "Sample type") +
  theme_bw() +
  theme(text = element_text(size = 13),
        legend.position = c(.8,.2)) +
    guides(color = "none", shape = "none", fill = "none")
 )

### Life mode

# plot top model predictions

# format data for lifemode plot
fit.newdata = expand.grid(method = c("S", "R"),
                          primer = c("COI", "16S"),
                                 size = 1)

fitsies <- fitted(life.prim.method, 
                  newdata = fit.newdata,probs = c(0.025, 0.1,
                                                  0.9, 0.975))

dim(fitsies)
fitsies # dims are expand grid rows (4), quantiles etc. (6), levels of Y (3)
names(fitsies[1,1,])

fitsies.list <- list()
for(i in 1:dim(fitsies)[3]){
  fitsies.list[[i]] <- fitsies[,,i]
}

life_fits_df <- do.call(rbind, fitsies.list) %>%
  as.data.frame() %>%
  mutate(method = rep(fit.newdata$method, 3),
         primer = rep(fit.newdata$primer, 3),
         life = c(rep("All larvae aquatic", 4),
                  rep("Fully terrestrial", 4),
                  rep("Some larvae aquatic", 4))) %>%
  group_by(life) %>%
  mutate(mean_p = mean(Estimate)) %>%
  ungroup() %>%
  mutate(life = fct_reorder(life, mean_p),
         method = case_when(method == "S" ~ "Understory",
                            method == "R" ~ "Bark"))

# create lifemode plot
(life_plot2 <-
life_fits_df %>%
    mutate(methodtreeprimer = paste0(method, " (", primer, ")")) %>%
  ggplot() +
  geom_errorbar(aes(y = life, xmin = 100*Q2.5, 
                    xmax = 100*Q97.5, color = methodtreeprimer,
                    group = methodtreeprimer),
                position = position_dodge(width = dod),
                width = 0, size = thin) +  
  geom_errorbar(aes(y = life, xmin = 100*Q10, 
                    xmax = 100*Q90, color = methodtreeprimer,
                    group = methodtreeprimer),
                size = thik,
                position = position_dodge(width = dod),
                width = 0) +
  geom_point(aes(y = life, x = 100*Estimate, color = methodtreeprimer,
                 fill = methodtreeprimer,
                 shape = methodtreeprimer),
             position = position_dodge(width = dod),
             size = psiz) +
  scale_color_manual(values = c("steelblue", "steelblue", "firebrick", "firebrick")) +
  scale_fill_manual(values = c("steelblue", "white", "firebrick", "white")) +
  scale_shape_manual(values = c(21,21,21,21)) +
  labs(color = "",
       y = "",
       x = "% of MOTUs",
       title = "Life mode",
       fill = "Sample type",
       shape = "Sample type",
       color = "Sample type") +
  theme_bw() +
  theme(text = element_text(size = 13),
        legend.position = c(.75,.15)) +
  guides(color = "none") +
  theme(legend.key.size = unit(.10, 'cm'), #change legend key size
        legend.key.height = unit(.10, 'cm'), #change legend key height
        legend.key.width = unit(.10, 'cm'), #change legend key width
        legend.title = element_text(size=9), #change legend title font size
        legend.text = element_text(size=9), #change legend text font size
        legend.background = element_rect(fill = "transparent"))  
 )

### Combine all 3 plots with patchwork
feed_plot2 / size_plot2 / life_plot2

# ggsave("figures/Figure5.jpg",
#        width = 5, height = 12, dpi = 400)
```
# Create Figure S1. 
Create Figure S1 in Allen et al. (2023). Read-based rarefaction curves.
```{r}

library(iNEXT)

# format data to create accumulation curves

### Zeale

# run accumulation curve analyses in iNEXT
# note: un-comment to run (takes a long time); or load pre-run version below
# accum.motu.reads.Z <- iNEXT(
#   as.data.frame(select(filter(art_motu_reads, grepl(id, pattern = "Z")), !contains(c("id", "coleop", "sum", "3CS")))),
#                             q=0, datatype="abundance", endpoint = 60000, knots = 100)
# saveRDS(accum.motu.reads.Z, "output/accum.motu.reads.Z_allMOTUs.rds")
accum.motu.reads.Z <- readRDS("output/accum.motu.reads.Z_allMOTUs.rds")
  
Z.df <- accum.motu.reads.Z$iNextEst$coverage_based

### Coleop

# run accumulation curve analyses in iNEXT
# note: un-comment to run (takes a long time); or load pre-run version below
# accum.motu.reads.C <- iNEXT(
#   as.data.frame(select(filter(art_motu_reads, grepl(id, pattern = "C")), !contains(c("id", "zeale", "sum")))),
#                             q=0, datatype="abundance", endpoint = 60000, knots = 100)
# saveRDS(accum.motu.reads.C, "output/accum.motu.reads.C_allMOTUs.rds")
accum.motu.reads.C <- readRDS("output/accum.motu.reads.C_allMOTUs.rds")

C.df <- accum.motu.reads.C$iNextEst$coverage_based

# make plots 

### MOTU level

### zeale 

(zrareplot <- Z.df %>%
  group_by(Assemblage) %>%
  mutate(pct = qD / max(qD)) %>%
  filter(Method %in% c("Rarefaction", "Observed")) %>%
ggplot() +
  geom_line(aes(x = m, y = qD, color = Assemblage)) +
  geom_line(aes(x = m, y = qD, color = Assemblage), data = filter(Z.df, Method == "Extrapolation"), lty = 2) +
  geom_point(aes(x = m, y = qD, color = Assemblage), data = filter(Z.df, Method == "Observed"), size = 2) +
  guides(color = "none") +
  scale_x_continuous(limits = c(0, 60000)) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x = "No. arthropod reads", y = "No. MOTUs (COI primer set)") +
  theme_bw() +
  theme(text = element_text(size = 12))
)
  
### coleop

(crareplot <- C.df %>%
  filter(Method %in% c("Rarefaction", "Observed")) %>%
ggplot() +
  geom_line(aes(x = m, y = qD, color = Assemblage)) +
  geom_line(aes(x = m, y = qD, color = Assemblage), data = filter(C.df, Method == "Extrapolation"), lty = 2) +
  geom_point(aes(x = m, y = qD, color = Assemblage), data = filter(C.df, Method == "Observed"), size = 2) +
  guides(color = "none") +
  scale_x_continuous(limits = c(0, 60000)) +
  scale_y_continuous(limits = c(0, 216)) +
  labs(x = "No. arthropod reads", y = "No. MOTUs (16S primer set)") +
  theme_bw() +
  theme(text = element_text(size = 12))
)

# combine into a single plot
gg <- gridExtra::arrangeGrob(crareplot, zrareplot, ncol = 2)
plot(gg)

# ggsave(plot = gg, "figures/FigureS1.png", dpi = 400, width = 6, height = 3.5)
```
# Create Figure S2. 
Create Figure S2 in Allen et al. (2023). Model and plot diversity, richness, and evenness by substrate, tree type, and primer set. Read data rarefied to 40000 reads (drops 8 samples).
```{r}

# MOTU rarefied species richness by sample and primer set
# format data
names(art_full) # to get column numbers of matrix
data_cols <- 9:45
coleop_rare_mat <- art_full %>%
  filter(primer == "coleop") %>%
  select(all_of(data_cols)) %>%
  as.matrix() %>%
  t() %>%
  as.data.frame() %>%
  tibble::rownames_to_column() %>%
  rowwise() %>% 
  mutate(sample.sum = sum(c_across(starts_with("V")), na.rm = T)) %>%
  as.data.frame()
rownames(coleop_rare_mat) <- coleop_rare_mat$rowname

zeale_rare_mat <- art_full %>%
  filter(primer == "zeale") %>%
  select(all_of(data_cols)) %>%
  as.matrix() %>%
  t() %>%
  as.data.frame() %>%
  tibble::rownames_to_column() %>%
  rowwise() %>% 
  mutate(sample.sum = sum(c_across(starts_with("V")), na.rm = T)) %>%
  as.data.frame()
rownames(zeale_rare_mat) <- zeale_rare_mat$rowname

# get list of samples with low read depth (< 40000)
lowreadsampsC40 <- coleop_rare_mat %>%
  filter(sample.sum < 40000); (lowreadsampsC40 <- lowreadsampsC40$rowname)

lowreadsampsZ40 <- zeale_rare_mat %>%
  filter(sample.sum < 40000); (lowreadsampsZ40 <- lowreadsampsZ40$rowname)

unique(c(lowreadsampsC40, lowreadsampsZ40))

# create versions of the dataset with only samples reaching threshold read depths

# 40000 threshold removes 8 samples ...
coleop_rare_mat40 <- coleop_rare_mat %>%
  filter(rowname %notin% unique(c(lowreadsampsC40, lowreadsampsZ40))) %>%
  select(-rowname)

zeale_rare_mat40 <- zeale_rare_mat %>%
  filter(rowname %notin% unique(c(lowreadsampsC40, lowreadsampsZ40))) %>%
  select(-rowname)

# get lowest number of reads
minreads40 <- min(c(coleop_rare_mat40$sample.sum, zeale_rare_mat40$sample.sum))

rarefy_coleop40 <-rarefy(select(coleop_rare_mat40,-sample.sum), 
                         minreads40, se = T)
rarefy_zeale40 <- rarefy(select(zeale_rare_mat40,-sample.sum), 
                         minreads40, se = T)

rarefy_coleop_df40 <- rarefy_coleop40 %>%
  t() %>%
  as.data.frame() %>%
  bind_cols(colnames(rarefy_coleop40),.) %>%
  rename(sample = 1) %>%
  mutate(primer = "16S")

rarefy_zeale_df40 <- rarefy_zeale40 %>%
  t() %>%
  as.data.frame() %>%
  bind_cols(colnames(rarefy_zeale40),.) %>%
  rename(sample = 1) %>%
  mutate(primer = "COI")

rarefy_df40 <- rarefy_coleop_df40 %>%
  bind_rows(rarefy_zeale_df40) %>%
  mutate(method = stringr::str_sub(sample, -1,-1),
         tree = stringr::str_sub(sample, -2,-2)) %>%
  mutate(method = case_when(method == "R" ~ "Bark",
                            TRUE ~ "Understory"),
         tree = case_when(tree == "D" ~ "Deciduous",
                          TRUE ~ "Coniferous"))

# now create a rarified version of the read matrixes for diversity metrics

coleop_rare_mat_rare40 <- rrarefy(select(coleop_rare_mat40,-sample.sum), 
                                  minreads40)

zeale_rare_mat_rare40 <- rrarefy(select(zeale_rare_mat40,-sample.sum), 
                                 minreads40)

# Calculate diversity metrics
coleop_rare_shannon40 <- diversity(coleop_rare_mat_rare40, 
                               index = "shannon", 
                               MARGIN = 1, base = exp(1)) %>%
  as.data.frame() %>%
  bind_cols(rownames(coleop_rare_mat_rare40),.) %>%
  rename(sample = 1, shannon = 2) %>%
  mutate(primer = "16S")

zeale_rare_shannon40 <- diversity(zeale_rare_mat_rare40, 
                               index = "shannon", 
                               MARGIN = 1, base = exp(1)) %>%
  as.data.frame() %>%
  bind_cols(rownames(coleop_rare_mat_rare40),.) %>%
  rename(sample = 1, shannon = 2) %>%
  mutate(primer = "COI")

shannon_df40 <- coleop_rare_shannon40 %>%
  bind_rows(zeale_rare_shannon40) %>%
  mutate(method = stringr::str_sub(sample, -1,-1),
         tree = stringr::str_sub(sample, -2,-2)) %>%
  mutate(method = case_when(method == "R" ~ "Bark",
                            TRUE ~ "Understory"),
         tree = case_when(tree == "D" ~ "Deciduous",
                          TRUE ~ "Coniferous"))

# Calculate Simpsons index on rarefied data
coleop_rare_simpson40 <- diversity(coleop_rare_mat_rare40, 
                               index = "simpson") %>%
  as.data.frame() %>%
  bind_cols(rownames(coleop_rare_mat_rare40),.) %>%
  rename(sample = 1, simpson = 2) %>%
  mutate(primer = "16S")

zeale_rare_simpson40 <- diversity(zeale_rare_mat_rare40, 
                               index = "simpson") %>%
  as.data.frame() %>%
  bind_cols(rownames(zeale_rare_mat_rare40),.) %>%
  rename(sample = 1, simpson = 2) %>%
  mutate(primer = "COI")

simpson_df40 <- coleop_rare_simpson40 %>%
  bind_rows(zeale_rare_simpson40) %>%
  mutate(method = stringr::str_sub(sample, -1,-1),
         tree = stringr::str_sub(sample, -2,-2)) %>%
  mutate(method = case_when(method == "R" ~ "Bark",
                            TRUE ~ "Understory"),
         tree = case_when(tree == "D" ~ "Deciduous",
                          TRUE ~ "Coniferous"))

# set font size for all plot annotations below
annfont <- 2
ytextsize = 10
xtextsize = 11

# Richness model

# un-comment to run the model, or load the pre-run version below 
library(brms)
# (Smod40 <- brm(log(S) ~ tree + method + primer, data = rarefy_df40, cores = 4,
#                iter = 4000, warmup = 2000))
# saveRDS(Smod40, "output/Smod40_swarm1_allMOTUs.rds")
(Smod40 <- readRDS("output/Smod_swarm1_allMOTUs.rds"))
# *** = 99% don't overlap; ** = 95% don't overlap; * = 90% don't overlap
#                  Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept            3.87      0.11     3.66     4.09 1.00    10164     6012
# treeDeciduous        0.21      0.12    -0.02     0.44 1.00     9926     6033*
# methodUnderstory     0.46      0.13     0.21     0.71 1.00     9559     5878***
# primerCOI           -0.82      0.12    -1.06    -0.59 1.00     9768     5750***

# check credible intervals
library(tidybayes)
quantile(spread_draws(Smod40, b_treeDeciduous)$b_treeDeciduous, c(0.05, 0.95)) 
# 0.01424768 0.40348055
quantile(spread_draws(Smod40, b_methodUnderstory)$b_methodUnderstory, c(0.005, 0.995)) 
# 0.1306643 0.8024770  
quantile(spread_draws(Smod40, b_primerCOI)$b_primerCOI, c(0.005, 0.995)) 
# -1.1305998 -0.4995662 

# format data for plot
Smod40_df <- conditional_effects(Smod40, prob = 0.95) %>%
  do.call(rbind, .) %>%
  mutate(var = c("Tree type", "Tree type", "Substrate", "Substrate", "Primer set", "Primer set"),
         effect1__ = case_when(effect1__ == "Coniferous" ~ "Conifer",
                              TRUE ~ effect1__)) %>%
  rename(q97.5 = upper__,
         q2.5 = lower__) %>%
  bind_cols(select(do.call(rbind, conditional_effects(Smod40, prob = 0.80)), q10 = lower__, q90 = upper__))

Smod_ann <- data.frame(var = c("Tree type", "Substrate", "Primer set"),
                       x = 1.5,
                       y = 2.25,
                       label = c("*", "***", "***"))

# create the plot
(Splot40 <- 
Smod40_df %>%
  ggplot() +
  geom_point(aes(x=effect1__, y = exp(estimate__)),
             size = 3) +
  geom_errorbar(aes(x = effect1__, ymin = exp(q2.5), ymax = exp(q97.5)), width = 0) +
  geom_errorbar(aes(x = effect1__, ymin = exp(q10), ymax = exp(q90)), width = 0, size = 1) +
  geom_text(aes(x = x, y = y, label = label), data = Smod_ann, size = 6) +
  facet_wrap(~var, scales = "free_x") +
  labs(y = "Rarefied MOTU richness", x = "") +
  theme_bw()
)

# Evenness model

# un-comment to run the model, or load the pre-run version below 

#  (Simpmod40 <- brm(simpson ~ tree + method + primer, data = simpson_df40, cores = 4,
#                  iter = 4000, warmup = 2000))
# saveRDS(Simpmod40, "output/Simpmod40_swarm1_allMOTUs.rds")
(Simpmod40 <- readRDS("output/Simpmod40_swarm1_allMOTUs.rds"))
#                  Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept            0.51      0.07     0.38     0.65 1.00     9502     5770
# treeDeciduous        0.09      0.07    -0.05     0.24 1.00     9794     6041
# methodUnderstory     0.12      0.08    -0.04     0.27 1.00     9329     5257
# primerCOI           -0.14      0.07    -0.28    -0.00 1.00    10668     6094*

# check credible intervals
library(tidybayes)
get_variables(Simpmod40)
quantile(spread_draws(Simpmod40, b_treeDeciduous)$b_treeDeciduous, c(0.05, 0.95)) 
# -0.02568263  0.21498163
quantile(spread_draws(Simpmod40, b_methodUnderstory)$b_methodUnderstory, c(0.05, 0.95))
# -0.01435372  0.24540480
quantile(spread_draws(Simpmod40, b_primerCOI)$b_primerCOI, c(0.05, 0.95))
# -0.26353715 -0.02569073 

# format data for the plot
Simpmod40_df <- conditional_effects(Simpmod40, prob = 0.95) %>%
  do.call(rbind, .) %>%
  mutate(var = c("Tree type", "Tree type", "Substrate", "Substrate", "Primer set", "Primer set"),
         effect1__ = case_when(effect1__ == "Coniferous" ~ "Conifer",
                              TRUE ~ effect1__)) %>%
  rename(q97.5 = upper__,
         q2.5 = lower__) %>%
  bind_cols(select(do.call(rbind, conditional_effects(Simpmod40, prob = 0.80)), q10 = lower__, q90 = upper__))

Simpmod_ann <- data.frame(var = c("Tree type", "Substrate", "Primer set"),
                       x = 1.5,
                       y = .15,
                       label = c("", "", "*"))

# create the plot
(Simpplot40 <-
Simpmod40_df %>%
  ggplot() +
  geom_point(aes(x=effect1__, y = estimate__),
             size = 3) +
  geom_errorbar(aes(x = effect1__, ymin = q2.5, ymax = q97.5), width = 0) +
  geom_errorbar(aes(x = effect1__, ymin = q10, ymax = q90), width = 0, size = 1) +
  geom_text(aes(x = x, y = y, label = label), data = Simpmod_ann, size = 6) +
  facet_wrap(~var, scales = "free_x") +
  labs(y = "Rarefied MOTU evenness (Simpson)", x = "") +
  theme_bw()
)
  
# Diversity model

# un-comment to run the model, or load the pre-run version below 

# (Shanmod40 <- brm(shannon ~ tree + method + primer, data = shannon_df40, cores = 4,
#                   iter = 4000, warmup = 2000))
# saveRDS(Shanmod40, "output/Shanmod40_swarm1_allMOTUs.rds")
(Shanmod40 <- readRDS("output/Shanmod40_swarm1_allMOTUs.rds"))
#                  Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept            1.26      0.20     0.87     1.67 1.00     9467     6051
# treeDeciduous        0.32      0.22    -0.11     0.75 1.00     9284     6324
# methodUnderstory     0.66      0.23     0.20     1.12 1.00     9069     5426***
# primerCOI           -0.44      0.22    -0.87    -0.02 1.00    10212     6177**

# check credible intervals
library(tidybayes)
get_variables(Shanmod40)
quantile(spread_draws(Shanmod40, b_treeDeciduous)$b_treeDeciduous, c(0.05, 0.95)) 
# -0.03117101  0.67701142 
quantile(spread_draws(Shanmod40, b_methodUnderstory)$b_methodUnderstory, c(0.995, 0.005))
# 1.27724991 0.03905679 
quantile(spread_draws(Shanmod40, b_primerCOI)$b_primerCOI, c(0.005, 0.995))
# -1.018687  0.119324  

# format data for the plot
Shanmod40_df <- conditional_effects(Shanmod40, prob = 0.95) %>%
  do.call(rbind, .) %>%
  mutate(var = c("Tree type", "Tree type", "Substrate", "Substrate", "Primer set", "Primer set"),
         effect1__ = case_when(effect1__ == "Coniferous" ~ "Conifer",
                              TRUE ~ effect1__)) %>%
  rename(q97.5 = upper__,
         q2.5 = lower__) %>%
  bind_cols(select(do.call(rbind, conditional_effects(Shanmod40, prob = 0.80)), q10 = lower__, q90 = upper__))

Shanmod_ann <- data.frame(var = c("Tree type", "Substrate", "Primer set"),
                       x = 1.5,
                       y = .1,
                       label = c("", "***", "**"))

# create the plot
(Shanplot40 <-
Shanmod40_df %>%
  ggplot() +
  geom_point(aes(x=effect1__, y = estimate__),
             size = 3) +
  geom_errorbar(aes(x = effect1__, ymin = q2.5, ymax = q97.5), width = 0) +
  geom_errorbar(aes(x = effect1__, ymin = q10, ymax = q90), width = 0, size = 1) +
  geom_text(aes(x = x, y = y, label = label), data = Shanmod_ann, size = 6) +
  facet_wrap(~var, scales = "free_x") +
  labs(y = "Rarefied MOTU diversity (Shannon)", x = "") +
  theme_bw()
)

# create combined plot of marginal effects for the 3 models
gg = gridExtra::arrangeGrob(Shanplot40, Splot40, Simpplot40)

plot(gg)

ggsave(plot = gg, filename = "figures/FigureS2.png", 
       height = 9, width = 4, dpi = 400)
```
# Create Figure S3
Create Figure S3 in Allen et al. (2023). Run PERMANOVA and NMDS analyses at MOTU level to evaluate community composition by substrate and tree type.
```{r}
# format data matrices
mds.data.motu.Z <- rbind(sprayZ_motu_mat,
                  rollZ_motu_mat)

mds.data.motu.C <- rbind(sprayC_motu_mat,
                  rollC_motu_mat)

# check for blank rows
sort(apply(mds.data.motu.Z, 1, max))
sort(apply(mds.data.motu.C, 1, max))

# remove blank rows (no blanks)
  # these result from samples where no MOTUs were detected

mds.data.motu.Z.nb <- mds.data.motu.Z
mds.data.motu.C.nb <- mds.data.motu.C

### MOTU level ###

# format covariate data - motu level
env.data.motu.Z <- data.frame(
  tree = str_sub(gsub(gsub(rownames(mds.data.motu.Z.nb), 
                           pattern="_zeale", replacement = ""),
                      pattern="_coleop", replacement = ""), 
                                      -2, -2),
  method = str_sub(gsub(gsub(rownames(mds.data.motu.Z.nb), 
                           pattern="_zeale", replacement = ""),
                      pattern="_coleop", replacement = ""), 
                                        -1, -1)) %>%
  mutate(tree = case_when(tree == "C" ~ "Conifer",
                          tree == "D" ~ "Decid."),
         method = case_when(method == "R" ~ "Bark",
                            method == "S" ~ "Leaves"))

### zeale 

# package vegan NMDS
# run the model by un-commenting, or load the pre-run version below
set.seed(110)
# arthNMDS.motu.Z = metaMDS(mds.data.motu.Z.nb,
#                      k=2, distance = 'jaccard', trymax = 10000)
# saveRDS(arthNMDS.motu.Z, "output/arthNMDS.motu.Z.swarm1_allMOTUs.rds")
arthNMDS.motu.Z <- readRDS("output/arthNMDS.motu.Z.swarm1_allMOTUs.rds")
arthNMDS.motu.Z # stress: 0.2396

plot1 <- ordiplot(arthNMDS.motu.Z, choices = c(1, 2))

sites.long1 <-
  BiodiversityR::sites.long(plot1, env.data = env.data.motu.Z) %>%
  mutate(axis1 = as.numeric(axis1),
         axis2 = as.numeric(axis2),
         Method = paste0(method, " - ", tree))
head(sites.long1)
str(sites.long1)

(
  plot.motu.Z <-
    ggplot(sites.long1) +
    geom_vline(
      xintercept = c(0),
      color = "grey70",
      linetype = 2
    ) +
    geom_hline(
      yintercept = c(0),
      color = "grey70",
      linetype = 2
    ) +
    scale_x_continuous(sec.axis = dup_axis(labels = NULL, name = NULL)) +
    scale_y_continuous(sec.axis = dup_axis(labels = NULL, name = NULL)) +
    scale_color_manual(values = c("steelblue", "steelblue", 
                                  "firebrick","firebrick" )) +
    scale_fill_manual(values = c("steelblue", "steelblue", 
                                  "firebrick","firebrick" )) +
    scale_shape_manual(values = c(24, 21, 24, 21)) +
    geom_point(
      aes(
        x = axis1,
        y = axis2,
        shape = Method,
        color = Method,
        fill = Method,
      ),
      size = 3,
      alpha = 0.85
    ) +
    annotate("text", x = -.5, y = -.5, label = "Stress = 0.240", hjust = 0, size = 3) +
    # coord_fixed(ratio = 1) +
    labs(
      y = "NMDS2",
      x = "NMDS1",
      color = "Substrate",
      shape = "Tree type"
    ) +
    theme_bw() +
    theme(text = element_text(size = 12)) +
    ggtitle("COI primer set") +
   guides(shape = "none", color = "none", fill = "none")
)


# Test for differences among Method and Tree type
# PERMANOVA
set.seed(12555)
perm_arth <- adonis2(mds.data.motu.Z.nb ~ method + tree, data = env.data.motu.Z, 
       permutations = 9999, method = "jaccard",
       strata = NULL, contr.unordered = "contr.sum",
       contr.ordered = "contr.poly")

perm_arth
#          Df SumOfSqs      R2      F Pr(>F)    
# method    1   1.1629 0.07404 2.7342 0.0001 ***
# tree      1   0.5091 0.03241 1.1970 0.0471 *  
# Residual 33  14.0358 0.89355                  
# Total    35  15.7079 1.00000 

### coleop 

# format environmental data
env.data.motu.C <- data.frame(
  tree = str_sub(gsub(gsub(rownames(mds.data.motu.C.nb), 
                           pattern="_zeale", replacement = ""),
                      pattern="_coleop", replacement = ""), 
                                      -2, -2),
  method = str_sub(gsub(gsub(rownames(mds.data.motu.C.nb), 
                           pattern="_zeale", replacement = ""),
                      pattern="_coleop", replacement = ""), 
                                        -1, -1)) %>%
  mutate(tree = case_when(tree == "C" ~ "Conifer",
                          tree == "D" ~ "Decid."),
         method = case_when(method == "R" ~ "Bark",
                            method == "S" ~ "Leaves"))

# package vegan NMDS
# run the model by un-commenting, or load the pre-run version below
set.seed(110)
# arthNMDS.motu.C = metaMDS(mds.data.motu.C.nb,
#                      k=2, distance = 'jaccard', trymax = 10000)
# saveRDS(arthNMDS.motu.C, "output/arthNMDS.motu.C.swarm1_allMOTUs.rds")
arthNMDS.motu.C <- readRDS("output/arthNMDS.motu.C.swarm1_allMOTUs.rds")
arthNMDS.motu.C # stress: 0.2126

plot1 <- ordiplot(arthNMDS.motu.C, choices = c(1, 2))

sites.long1 <-
  BiodiversityR::sites.long(plot1, env.data = env.data.motu.C) %>%
  mutate(axis1 = as.numeric(axis1),
         axis2 = as.numeric(axis2),
         Method = paste0(method, " - ", tree))
head(sites.long1)
str(sites.long1)

(
  plot.motu.C <-
    ggplot(sites.long1) +
    geom_vline(
      xintercept = c(0),
      color = "grey70",
      linetype = 2
    ) +
    geom_hline(
      yintercept = c(0),
      color = "grey70",
      linetype = 2
    ) +
    scale_x_continuous(sec.axis = dup_axis(labels = NULL, name = NULL)) +
    scale_y_continuous(sec.axis = dup_axis(labels = NULL, name = NULL)) +
    scale_color_manual(values = c("steelblue", "steelblue", 
                                  "firebrick","firebrick" )) +
    scale_fill_manual(values = c("steelblue", "steelblue", 
                                  "firebrick","firebrick" )) +
    scale_shape_manual(values = c(24, 21, 24, 21)) +
    geom_point(
      aes(
        x = axis1,
        y = axis2,
        shape = Method,
        color = Method,
        fill = Method,
      ),
      size = 3,
      alpha = 0.85
    ) +
    annotate("text", x = -.5, y = -.5, label = "Stress = 0.213", hjust = 0, size = 3) +
    # coord_fixed(ratio = 1) +
    labs(
      y = "NMDS2",
      x = "NMDS1",
      color = "Substrate",
      shape = "Tree type"
    ) +
    theme_bw() +
    theme(text = element_text(size = 12)) +
    ggtitle("16S primer set") +
   guides(shape = "none", color = "none", fill = "none")
)


# Test for differences among Method and Tree type
# PERMANOVA
set.seed(12555)
perm_arth <- adonis2(mds.data.motu.C.nb ~ method + tree, data = env.data.motu.C, 
       permutations = 9999, method = "jaccard",
       strata = NULL, contr.unordered = "contr.sum",
       contr.ordered = "contr.poly")

perm_arth
#          Df SumOfSqs      R2      F Pr(>F)    
# method    1   1.2115 0.08100 3.1163 0.0001 ***
# tree      1   0.5269 0.03523 1.3555 0.0111 *  
# Residual 34  13.2173 0.88376                  
# Total    36  14.9557 1.00000     

# combine into a single plot
gg <- gridExtra::arrangeGrob(plot.motu.Z, plot.motu.C, ncol = 2)
plot(gg)

# ggsave(plot = gg, "figures/FigureS3.png", dpi = 400, width = 6, height = 3.5)
```
# Create Tables S1 and S2
These were created manually using data summarized in the script "pinebugs_post_process_final.Rmd".
# Create Table S3. 
Create Table S3 in Allen et al. (2023). Non-Arthropod taxa identified using the COI and 16S primer sets.
```{r}
mamgen <- c("Blarina", "Canis", "Didelphis", "Eptesicus", 
"Felis", "Glaucomys", "Lasiurus", "Mephitis", "Odocoileus", 
"Peromyscus", "Procyon", "Rattus", "Sciurus", 
"Sylvilagus", "Tamias", "Vulpes")

farm_gen <- c("Bos", "Capra", "Equus", "Ovis", "Sus")

amphib_gen <- c("Hemidactylium", "Plethodon")

# make dataframe of non-arthropod taxa

### 16S Coleop

c_other <- read.csv("data/pinebugs.samples.step2.negsub.swarm1.C_allMOTUs.csv") %>%
  filter(phylum != "Arthropoda",
         !is.na(phylum)) %>%
  mutate(label = case_when(genus %in% mamgen ~ "Chordata (wild mammals)",
                   family == "Sciuridae" ~ "Chordata (wild mammals)",
                   genus %in% farm_gen ~ "Chordata (farm animals)",
                   genus %in% amphib_gen ~ "Chordata (amphibians)",
                   family == "Corvidae" ~ "Chordata (birds)",
                   phylum == "Basidiomycota" ~ "Basidiomycota",
                   phylum == "Ascomycota" ~ "Ascomycota",
                   phylum == "Mollusca" ~ "Mollusca",
                   phylum == "Annelida" ~ "Annelida",
                   TRUE ~ "Chordata"
                   )) %>%
  group_by(label) %>%
  summarize(across(contains("sample"), sum),
            .groups = "drop") %>%
  rowwise() %>%
  mutate(primer16S = sum(c_across(contains("sample")))) %>%
  ungroup() %>%
  select(label, primer16S)

### COI Zeale

z_other <- read.csv("data/pinebugs.samples.step2.negsub.swarm1.Z_allMOTUs.csv") %>%
  filter(phylum != "Arthropoda",
         !is.na(phylum)) %>%
  group_by(phylum) %>%
  summarize(across(contains("sample"), sum),
            .groups = "drop") %>%
  rowwise() %>%
  mutate(COI = sum(c_across(contains("sample")))) %>%
  ungroup() %>%
  select(label = phylum, COI)

### Combine the 16S and COI data frames into one

both_other <- c_other %>%
  full_join(z_other)

# a list of the "wild" (free ranging) mammals for reference
mamlist <- c("Blarina brevicauda", "Canis lupus", "Didelphis virginiana", 
"Eptesicus fuscus", "Felis catus", "Glaucomys volans", "Lasiurus borealis", 
"Mephitis mephitis", "Odocoileus virginianus", "Peromyscus leucopus", 
"Procyon lotor", "Rattus norvegicus", "Sciurus carolinensis", 
"Sylvilagus floridanus", "Tamias striatus", "Vulpes vulpes"
)
length(mamlist)

# view the table
both_other

# export the table
library(knitr)
library(kableExtra)
library(magrittr)

kable(both_other, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  cat(., file = "figures/TableS3.html")

```
# Create Table S4
Create Table S4 in Allen et al. (2023). Total reads by order, family, genus, species, and primer set.
```{r}
# get family common names
# un-comment to run this script; otherwise read in the family common name data below
# common_fam <- taxize::sci2comm(sort(unique(art_full$family)))
# commons_df <- common_fam %>%
#   do.call(rbind, .) %>%
#   as.data.frame() %>%
#   rename(common_fam = 1) %>%
#   arrange(common_fam)
# commons_df$family <- rownames(commons_df)
# write.csv(commons_df, "data/family_commons.csv", row.names = F)
(commons_df <- read.csv("data/family_commons.csv"))

# format
master <- art_full  %>%
  mutate(engname = case_when(order == "Mesostigmata" ~ "Mites",
                          order == "Sarcoptiformes" ~ "Mites",
                          order == "Trombidiformes" ~ "Mites",
                          class == "Collembola" ~ "Springtails",
                          class == "Chilopoda" ~ "Centipedes",
                          order == "Araneae" ~ "Spiders",
                          order == "Opiliones" ~ "Harvestmen",
                          order == "Blattodea" ~ "Cockroaches",
                          order == "Coleoptera" ~ "Beetles",
                          order == "Dermaptera" ~ "Earwigs",
                          order == "Diptera" ~ "Flies",
                          order == "Ephemeroptera" ~ "Mayflies",
                          order == "Hemiptera" ~ "True bugs",
                          order == "Hymenoptera" ~ "Bees, wasps, & ants",
                          order == "Lepidoptera" ~ "Butterflies & moths",
                          order == "Mantodea" ~ "Mantids",
                          order == "Megaloptera" ~ "Dobsonflies",
                          order == "Cyclopoida" ~ "Copepods",
                          order == "Neuroptera" ~ "Lacewings",
                          order == "Odonata" ~ "Dragonflies & damselflies",
                          order == "Orthoptera" ~ "Crickets & grasshoppers",
                          order == "Phasmatodea" ~ "Walking sticks",
                          order == "Phthiraptera" ~ "Thrips",
                          order == "Psocoptera" ~ "Bark lice",
                          order == "Siphonaptera" ~ "Fleas",
                          order == "Trichoptera" ~ "Caddisflies",
                          (is.na(order) & is.na(class) & phylum == "Arthropoda") ~ "Arthropod sp.",
                          (is.na(order) & !is.na(class) ~ paste0(class, " sp."))
                          )) %>%
  mutate(lowestname = case_when(!is.na(species) ~ species,
                           !is.na(genus) & is.na(species) ~ paste0(genus, " sp."),
                           is.na(genus) ~ NA)) %>%
  left_join(commons_df) # add common family name

# format data for spray samples
orderreads_S4 <- master %>%
  group_by(engname, phylum, class, order, family, genus, species, lowestname, common_fam, primer) %>%
  summarize(across(contains(c("CS", "DS", "CR", "DR")), sum),
            .groups = "drop") %>%
  rowwise() %>% 
  mutate(sum = sum(c_across(contains(c("CS", "DS", "CR", "DR"))), na.rm = T)) %>%
  ungroup() %>%
  mutate(order2 = case_when(!is.na(order) ~ paste0(order, " (", engname, ")"),
                            TRUE ~ NA),
         family2 = case_when(!is.na(family) & !is.na(common_fam) ~ paste0(family, " (", common_fam, ")"),
                             !is.na(family) & is.na(common_fam) ~ family,
                             TRUE ~ NA)) %>%
  select(class, order2, family2, lowestname, primer, sum) %>%
  arrange(class, order2, family2, lowestname) %>%
  pivot_wider(id_cols = class:lowestname, names_from = primer, values_from = sum) %>%
  replace_na(list(zeale = 0, coleop = 0)) %>%
  # mutate(class = "") %>%
  rename(Class = class, Order = order2, Family = family2, Species = lowestname, primerCOI = zeale, primer16S = coleop)

# view the table
orderreads_S4

# export the table
library(knitr)
library(kableExtra)
library(magrittr)

kable(orderreads_S4, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  cat(., file = "figures/TableS4.html")

```
# Examining reference databases
```{r}

library(data.table) # data.table_1.14.8

# load the 16S reference data set, sub-set Arthropods, and count sequences
cref <- fread("data/ncbi.bold.art.hum.mus.may23.pcr.pga.tax.uni.cln.C.tsv") %>%
  mutate(len = nchar(sequence)) %>%
  filter(phylum == "Arthropoda")
length(unique(cref$species)) # 33540 species
length(unique(cref$order)) # 92 orders
length(unique(cref$family)) # 1324 families
nrow(cref) # 44181 sequences

# load the COI reference data set, sub-set Arthropods, and count sequences
zref <- fread("data/ncbi.bold.art.hum.mus.may23.pcr.pga.tax.uni.cln.Z.tsv") %>%
  mutate(len = nchar(sequence)) %>%
  filter(phylum == "Arthropoda")

length(unique(zref$species)) # 58323 species
length(unique(zref$order)) # 90 orders 
length(unique(zref$family)) # 1172 families
nrow(zref) # 355000 sequences
```
# Random calculations for the results and discussion
```{r}

# how many carabids in COI samples
sum(apply(select(art_full[!is.na(art_full$family) & art_full$family=="Carabidae",], contains(c("CS", "CR", "DS", "DR"))), 2, sum)>0)
# 31

# out of
length(apply(select(art_full, contains(c("CS", "CR", "DS", "DR"))), 2, sum)>0)
# 37

# how many MOTUs not identified to species
(cs3 %>% filter(!is.na(species)) %>% tally())
nrow(cs3)
117/1077 # 0.1086

(zs3 %>% filter(!is.na(species)) %>% tally())
nrow(zs3)
228/650 # 0.3519

# how many insect species and families (no spiders etc.)
cs3insect <- cs3 %>% 
  filter(class == "Insecta")

zs3insect <- zs3 %>% 
  filter(class == "Insecta")

unique(c(cs3insect$family, zs3insect$family)) # 159 families
unique(c(cs3insect$species, zs3insect$species)) # 278 species

# time estimates from Anthony and Bobby
# 
# Extractions (20 hrs)
# PCR optimization (40-80 hrs)
# Inhibitor Cleanup (4 hrs)
# PCR1 (each primer set) (16 hrs)
# Bead 1 (6 hrs)
# Gels (6 hrs)
# Quibit + Pool (4 hrs)
# PCR 2 (8 hrs)
# Bead 2- 4 hrs
# Qubit 2 (4 hrs)
# Gel 2 (4 hrs)
# Pool/ send for sequencing (3 hrs)

20+40+4+16+6+6+4+8+4+4+4+3
20+80+4+16+6+6+4+8+4+4+4+3
```


