# Post-processing of arthropod metabarcoding data surface eDNA samples
40 eDNA samples were taken from tree bark ('roller aggregation') and understory foliage ('spray aggregation') in Colliers Mills Wildlife Management Area, New Jersey, USA and processed using COI and 16S metabarcoding. This code allows reproduction of bioinformatics post-processing steps for the paper Allen et al. (2023).

Allen, M. C., Lockwood, J. L., Kwait, R., Vastano, A., Peterson, D. L. Tkacenko, L. A., … Jaffe, B. D. (2023). Using surface environmental DNA to assess arthropod biodiversity within a forested ecosystem. Environmental DNA.

The code takes the raw ecotag output and the raw MOTU table, joins them, and performs all filtering steps to get to the final data set. Note that there is a break in the code where sequences were exported to a fasta file (make_blasta.R), searched using BLAST on the cluster (custom bash script), and then re-combined with the ecotag annotated MOTU table for manual curation. This code also allows you to skip that labor intensive step by reading in the final MOTU table with manually curated taxonomy (see Methods in Allen et al. 2023).

# read in data and load libraries
```{r}
library(dplyr) # dplyr_1.1.2
library(data.table) # data.table_1.14.8
'%notin%' <- Negate('%in%')

### Zeale ###

# Read in Zeale (COI) ecotag results 
z <- fread("data/merged.uni.c10.l130.L185.sht.srt.nochi.1line.swarm1.fix.tag.ann.srt.Z.tsv") %>%
  rename(match = 3) # 819 MOTUs

# 138 are well matched to species already, so 681 to BLAST
test <- z %>% filter(match >= 0.99,
             !is.na(species_name))
nrow(test)

# 123 unique species names (before BLAST and manual curation)
sort(unique(test$scientific_name))
rm(test)

# that leaves 681 to blast if desired (to format fasta for this, see next section)
zblast <- z %>% 
  filter(!(match >= 0.99 & !is.na(species_name)))

# Read in Zeale (COI) primer read counts by sample (MOTU table)
zc <- fread("data/merged.uni.c10.l130.L185.sht.srt.nochi.1line.swarm1.Z_output.counts.csv")

zz <- z %>%
  left_join(zc, by = join_by(id, definition, sequence))

### Coleop ###

# Read in Coleop (16S) ecotag results
c <- fread("data/merged.uni.c10.l75.L125.sht.srt.nochi.1line.swarm1.fix.tag.ann.srt.C.tsv") %>%
  rename(match = 3) # 1179 MOTUs

# 116 are well matched to species already
test2 <- c %>% filter(match >= 0.99,
             !is.na(species_name))
nrow(test2)

# 116 unique species names (before BLAST and manual curation)
sort(unique(test2$scientific_name))
rm(test2)

# that leaves 1063 to blast
cblast <- c %>% 
  filter(!(match >= 0.99 & !is.na(species_name)))

# Read in Zeale (COI) primer read counts by sample
c2 <- fread("data/merged.uni.c10.l75.L125.sht.srt.nochi.1line.swarm1.C_output.counts.csv")

cc <- c %>%
  left_join(c2, by = join_by(id, definition, sequence))

```
# make fasta files to cross-check identifications using BLAST searches (if applicable)
Here, I used code in the process_blast.Rmd script to create a fasta file of sequences to BLAST as a cross-check on the ecotag results. To do that, run the code up until here, then open and run process_blast.Rmd. To repeat exactly what I did, you'd need to then perform blastn searches on the fasta file (I did it on the Rutgers Amarel cluster) and then collect and rejoin the output to the ecotag results using process_blast.Rmd. Then do manual curation of the resulting csv file to pick the most appropriate final identification for each MOTU (see Methods of Allen et al. 2023). Or, you can skip ahead to the results of that effort in the "separate the samples and negatives" section. 
# read in the  manually curated MOTU identifications & join to read count data
At this point, the identifications were manually curated (reivewed against BLAST matches and GBIF range maps) and the csv file (e.g., "zzblastID_swarm1_finaltax.csv") had a complete column named overrideID. That column contained the appropriate NCBI taxon name in cases where BLAST or geography caused a manual override of the ecotag ID. The script below selects the appropriate final ID (ecotagID or overrideID) and puts it in a new column called finaltax. That column was then used to look up the higher-level taxonomy using the taxize package. Note that 'finaltax' was not necessarily at an appropriate taxonomic level given the strength of the match (e.g., it may have been ID'd to species by ecotag with only a 95% match). Later, the taxonomic resolution will be manually adjusted where appropriate based on defined match thresholds.
```{r}

### Zeale ###

zzblastID.cur <- read.csv("data/zzblastID_swarm1_finaltax_allMOTUs.csv")

zz_fin <- zz %>%
  left_join(select(zzblastID.cur, id, bestblastmatch, overrideID, blastnotes)) %>%
  select(id, ecotagmatch = match, bestblastmatch, ecotagID = scientific_name, overrideID, 
         count = total_reads, wt = cluster_weight, starts_with("sample"), sequence, blastnotes) %>%
  mutate(finaltax = case_when(overrideID == "" ~ ecotagID,
                              is.na(overrideID) ~ ecotagID,
                              TRUE ~ overrideID),
         bestmatch = case_when(ecotagmatch >= bestblastmatch ~ ecotagmatch,
                               bestblastmatch > ecotagmatch ~ bestblastmatch,
                               is.na(bestblastmatch) ~ ecotagmatch))

### Coleop ###

ccblastID.cur <- read.csv("data/ccblastID_swarm1_finaltax_allMOTUs.csv")

cc_fin <- cc %>%
  left_join(select(ccblastID.cur, id, bestblastmatch, overrideID, blastnotes)) %>%
  select(id, ecotagmatch = match, bestblastmatch, ecotagID = scientific_name, overrideID, 
         count = total_reads, wt = cluster_weight, starts_with("sample"), sequence, blastnotes) %>%
  mutate(finaltax = case_when(overrideID == "" ~ ecotagID,
                              is.na(overrideID) ~ ecotagID,
                              TRUE ~ overrideID),
         bestmatch = case_when(ecotagmatch >= bestblastmatch ~ ecotagmatch,
                               bestblastmatch > ecotagmatch ~ bestblastmatch,
                               is.na(bestblastmatch) ~ ecotagmatch))

```
# Add higher taxa & adjust taxonomy based on match score
Also separate the samples and negatives. full_taxa_lookup.csv was created using the package taxize in scripts/add_higher_taxa.R.
```{r}

# load look up table for higher taxa
full_taxa_lookup <- read.csv("data/full_tax_lookup.csv") %>%
  distinct() %>%
  mutate(class = case_when(order == "Malawimonadida" ~ "Malawimonadida",
                           TRUE ~ class),
         phylum = case_when(order == "Malawimonadida" ~ "Malawimonadida",
                            TRUE ~ phylum))

# check that taxa are in the lookup table
alltaxcur <- unique(c(cc_fin$finaltax, zz_fin$finaltax)) 
dput(alltaxcur[alltaxcur %notin% full_taxa_lookup$finaltax]) # character(0) = they are all there

#### Zeale ####

# add final taxonomic IDs & adjust based on match criterion

zz_finID <- zz_fin  %>%
  left_join(full_taxa_lookup, by = join_by(finaltax)) %>%
  select(id, finaltax, ecotagID, ecotagmatch, bestblastmatch, bestmatch, phylum, class, order, family, genus, species,
         count, wt, starts_with("sample"), sequence, blastnotes) %>%
  mutate(genus = case_when(bestmatch < 0.98 ~ NA,
                           TRUE ~ genus),
         species = case_when(bestmatch < 0.99 ~ NA,
                             TRUE ~ species),
         family = case_when(bestmatch < 0.95 ~ NA,
                            TRUE ~ family),
         order = case_when(bestmatch < 0.9 ~ NA,
                           TRUE ~ order),
         finalID = case_when(is.na(species) & is.na(genus) & is.na(family) & is.na(order) & is.na(class) ~ phylum,
                                    is.na(species) & is.na(genus) & is.na(family) & is.na(order) & !is.na(class) ~ class,
                                    is.na(species) & is.na(genus) & is.na(family) & !is.na(order) ~ order,
                                    is.na(species) & is.na(genus) & !is.na(family) ~ family,
                                    is.na(species) & !is.na(genus) ~ genus,
                                    !is.na(species) ~ species)) %>%
  select(id, finalID, bestmatch, phylum, class, order, family, genus, species,
         count, wt, starts_with("sample"), sequence) 

# check which columns are negatives
colnames(zz_finID) # input those column numbers in the next line
negatives <- c(12,13,21,51:53)
colnames(as.data.frame(zz_finID)[, negatives])

# select out just the samples
zs <- zz_finID %>%
  select(-all_of(negatives))

# sum the reads across all samples
sum(select(zs, starts_with("sample"))) # 4538107

# select out just the negatives
zn <- zz_finID %>%
  select(id, finalID, bestmatch, phylum, class, order, family, genus, species,
         count, wt, all_of(negatives), sequence)

# sum the reads across all negatives
sum(select(zn, starts_with("sample"))) # 1248

# get the max read count of each seqid present in the negatives
colnames(zn) # check which columns contain negative samples
zn$seqmax <- apply(zn[,12:17],1, max)

# write.csv(zs, file = "data/pinebugs.samples.step0.nofilters.swarm1.Z_allMOTUs.csv", row.names = F)
# write.csv(zn, file = "data/pinebugs.negatives.step0.nofilters.swarm1.Z_allMOTUs.csv", row.names = F)

#### Coleop ####

# add final taxonomic IDs

cc_finID <- cc_fin  %>%
  left_join(full_taxa_lookup, by = join_by(finaltax)) %>%
  select(id, finaltax, ecotagmatch, bestblastmatch, bestmatch, phylum, class, order, family, genus, species,
         count, wt, starts_with("sample"), sequence, blastnotes) %>%
  mutate(genus = case_when(bestmatch < 0.98 ~ NA,
                           TRUE ~ genus),
         species = case_when(bestmatch < 0.99 ~ NA,
                             TRUE ~ species),
         family = case_when(bestmatch < 0.95 ~ NA,
                            TRUE ~ family),
         order = case_when(bestmatch < 0.9 ~ NA,
                           TRUE ~ order),
         finalID = case_when(is.na(species) & is.na(genus) & is.na(family) & is.na(order) & is.na(class) ~ phylum,
                                    is.na(species) & is.na(genus) & is.na(family) & is.na(order) & !is.na(class) ~ class,
                                    is.na(species) & is.na(genus) & is.na(family) & !is.na(order) ~ order,
                                    is.na(species) & is.na(genus) & !is.na(family) ~ family,
                                    is.na(species) & !is.na(genus) ~ genus,
                                    !is.na(species) ~ species)) %>%
  select(id, finalID, bestmatch, phylum, class, order, family, genus, species,
         count, wt, starts_with("sample"), sequence) 

# check which columns are negatives
colnames(cc_finID) # input those column numbers in the next line
negatives.C <- c(12,13,21,22,53:65)
colnames(as.data.frame(cc_finID)[,negatives.C])

# select out just the samples
cs <- cc_finID %>%
  select(-all_of(negatives.C))

# sum the reads across all samples
sum(select(cs, starts_with("sample"))) # 2668687

# select out just the negatives
cn <- cc_finID %>%
  select(id, finalID, bestmatch, phylum, class, order, family, genus, species,
         count, wt, all_of(negatives.C), sequence)

# sum the reads across all negatives
sum(select(cn, starts_with("sample"))) # 60072

# get the max read count of each seqid present in the negatives
colnames(cn) # check which columns contain negative samples
cn$seqmax <- apply(cn[,12:17],1, max)

# write.csv(cs, file = "data/pinebugs.samples.step0.nofilters.swarm1.C_allMOTUs.csv", row.names = F)
# write.csv(cn, file = "data/pinebugs.negatives.step0.nofilters.swarm1.C_allMOTUs.csv", row.names = F)
```
# remove humans & mice
```{r}

### Coleop

cs1 <- cs %>%
  filter(genus %notin% c("Mus", "Homo"))

# sum the reads across all samples
sum(select(cs1, starts_with("sample"))) # 2611101

cn1 <- cn %>%
  filter(genus %notin% c("Mus", "Homo"))

# sum the reads across all samples
sum(select(cn1, starts_with("sample"))) # 6413

### Zeale

zs1 <- zs %>%
  filter(genus %notin% c("Mus", "Homo"))

# sum the reads across all samples
sum(select(zs1, starts_with("sample"))) # 4538107

zn1 <- zn %>%
  filter(genus %notin% c("Mus", "Homo"))

# sum the reads across all samples
sum(select(zn1, starts_with("sample"))) # 1248

```

# subtract contamination found in negative controls
```{r}

#### Zeale ####

# add the max reads by sequence in all the negatives
zs2a <- zs1 %>%
  left_join(select(zn1, id, seqmax), by = "id") %>%
  data.frame()
zs2 <- zs2a

colnames(zs2) # 12:47 are the samples

# zero out samples with less reads of a sequence than the max contamination among negative of that sequence
for(i in 12:47){
  tmp <- zs2a[,i] - zs2a$seqmax
  tmp2 <- as.numeric(ifelse(tmp<0,0,tmp))
  zs2[,i] <- tmp2
}

# count remaining reads in samples
sum(select(zs2, starts_with("sample"))) # 4534342
rm(zs2a)

# look at sequence lengths
hist(nchar(zs2$sequence))
# zs2$len <- nchar(zs2$sequence)

# write.csv(zs2, "data/pinebugs.samples.step2.negsub.swarm1.Z_allMOTUs.csv")

#### Coleop ####

# add the max reads by sequence in all the negatives
cs2a <- cs1 %>%
  left_join(select(cn1, id, seqmax), by = "id") %>%
  data.frame()
cs2 <- cs2a

colnames(cs2) # 12:48 are the samples

# zero out samples with less reads of a sequence than the max contamination among negative of that sequence
for(i in 12:48){
  tmp <- cs2a[,i] - cs2a$seqmax
  tmp2 <- as.numeric(ifelse(tmp<0,0,tmp))
  cs2[,i] <- tmp2
}

# count remaining reads in samples
sum(select(cs2, starts_with("sample"))) # 2603965
rm(cs2a)

# look at sequence lengths
hist(nchar(cs2$sequence))
# cs2$len <- nchar(cs2$sequence)

# write.csv(cs2, "data/pinebugs.samples.step2.negsub.swarm1.C_allMOTUs.csv")

```
# remove non-arthropods
```{r}

### Zeale ###

# check on phylum list and look at MOTUs with no phylum
unique(zs2$phylum)
zs2[is.na(zs2$phylum), c("id", "finalID")] # should be only MOTUs that were best matched to some taxon above phylum

# filter out non-arthropod - samples
zs3 <- zs2 %>%
  filter(phylum == "Arthropoda")

# add a column tallying reads within just the samples
names(zs3)
zs3$sampsum <- apply(zs3[,12:47], 1, sum)

# sum the reads across all samples
sum(select(zs3, starts_with("sample"))) # 4117138

# filter out non-arthropod - negatives
zn3 <- zn1 %>%
  filter(phylum == "Arthropoda")

# add a column tallying reads within just the negatives
names(zn3)
zn3$negsum <- apply(zn3[,12:17], 1, sum)

# sum the reads across all negatives
sum(select(zn3, starts_with("sample"))) # 997

# write.csv(zs3, file = "data/pinebugs.samples.onlyarth.negsub.Z_allMOTUs.csv", row.names = F)
# write.csv(zn3, file = "data/pinebugs.negatives.onlyarth.Z_allMOTUs.csv", row.names = F)

### Coleop ###

# check on phylum list and look at MOTUs with no phylum
unique(cs2$phylum)
cs2[is.na(cs2$phylum), c("id", "finalID")] # should be only MOTUs that were best matched to some taxon above phylum

# filter out non-arthropod - samples
cs3 <- cs2 %>%
  filter(phylum == "Arthropoda",
         id != "C_871") # remove the known contaminant species: Coconut Rhinocerous Beetle

# add a column tallying reads within just the samples
names(cs3)
cs3$sampsum <- apply(cs3[,12:48], 1, sum)

# sum the reads across all samples
sum(select(cs3, starts_with("sample"))) # 2240233

# filter out non-arthropods - negatives
cn3 <- cn1 %>%
  filter(phylum == "Arthropoda",
         id != "C_871") # remove the known contaminant species: Coconut Rhinocerous Beetle


# add a column tallying reads within just the negatives
names(cn3)
cn3$negsum <- apply(cn3[,12:28], 1, sum)

# sum the reads across all negatives
sum(select(cn3, starts_with("sample"))) # 4442

# write.csv(cs3, file = "data/pinebugs.samples.onlyarth.negsub.C_allMOTUs.csv", row.names = F)
# write.csv(cn3, file = "data/pinebugs.negatives.onlyarth.C_allMOTUs.csv", row.names = F)

```
# tallying orders, families, species
```{r}

### Zeale ###

# what species; n = 197
zs3 %>% filter(!is.na(species),
               sampsum > 0) %>% 
  select(species) %>%
  arrange(species) %>%
  distinct()

# what genus; n = 227
zs3 %>% filter(!is.na(genus),
               sampsum > 0) %>% 
  select(genus) %>%
  arrange(genus) %>%
  distinct()

# what families; n = 117 
zs3 %>% filter(!is.na(family),
               sampsum > 0) %>% 
  select(family) %>%
  arrange(family) %>%
  distinct()

# what orders; n = 16
zs3 %>% filter(!is.na(order)) %>% 
  select(order) %>%
  arrange(order) %>%
  distinct()

### Coleop ###

# note: no change to species, genera, families, or orders by examining MOTUs with match < 80%

# what species; n = 115
cs3 %>% filter(!is.na(species),
               sampsum > 0) %>% 
  select(species) %>%
  arrange(species) %>%
  distinct()

# what genus; n = 173
cs3 %>% filter(!is.na(genus),
               sampsum > 0) %>% 
  select(genus) %>%
  arrange(genus) %>%
  distinct()

# what families; n = 127
cs3 %>% filter(!is.na(family),
               sampsum > 0) %>% 
  select(family) %>%
  arrange(family) %>%
  distinct()

# what orders; n = 23
cs3 %>% filter(!is.na(order),
               sampsum > 0) %>% 
  select(order) %>%
  arrange(order) %>%
  distinct()

### combined 16S & COI

unique(art_order_motus$order); length(unique(art_order_motus$order))
unique(art_fam_motus$family); length(unique(art_fam_motus$family))
unique(art_gen_motus$genus); length(unique(art_gen_motus$genus))
unique(art_sp_reads$species); length(unique(art_sp_reads$species))

```
# counting reads for supplemental tables S1 & S2
```{r}

### Zeale

# Zeale step 1: how many reads after filtering steps performed on cluster

step1 <- z %>%
  left_join(zc, by = join_by(id, definition, sequence))

# check which columns are negatives
colnames(step1) # input those column numbers in the next line
negatives <- c(16,17,25,55:57)
colnames(as.data.frame(step1)[, negatives])

# select out just the samples
step1.s <- step1 %>%
  select(-all_of(negatives))

# tally up read stats for samples
sum(apply(select(step1.s, starts_with("sample")), 2, sum)) # 4538107
mean(apply(select(step1.s, starts_with("sample")), 2, sum)) # 126058.5
sd(apply(select(step1.s, starts_with("sample")), 2, sum)) # 38706.53
length(apply(select(step1.s, starts_with("sample")), 2, sum)) # 36
38706.53 / sqrt(36) # 6451

# select out just the negatives
step1.n <- step1 %>%
  select(id, all_of(negatives), sequence)

# tally up read stats for negatives
sum(apply(select(step1.n, starts_with("sample")), 2, sum)) # 1248
mean(apply(select(step1.n, starts_with("sample")), 2, sum)) # 208
sd(apply(select(step1.n, starts_with("sample")), 2, sum)) # 418.8
length(apply(select(step1.n, starts_with("sample")), 2, sum)) # 6
418.8 / sqrt(6) # 171.0

# Zeale step 2: remove human and mouse 

step2.s <- step1.s %>%
  left_join(select(zs, id:species)) %>%
  select(id, match, finalID:species, starts_with("sample")) %>%
  filter(species %notin% c("Homo sapiens", "Mus musculus")) # no mice or humans in COI primer

# Zeale step 3: subtract negatives, remove rest of non-arthropods
  # this is same as zs3 from above
step3.s <- zs3

# tally up read stats for samples
sum(apply(select(step3.s, starts_with("sample")), 2, sum)) # 4117138
mean(apply(select(step3.s, starts_with("sample")), 2, sum)) # 114364.9
sd(apply(select(step3.s, starts_with("sample")), 2, sum)) # 38707.6
length(apply(select(step3.s, starts_with("sample")), 2, sum)) # 36
38707.6 / sqrt(36) # 6451.267

step3.n <- zn3

# tally up read stats for samples
sum(apply(select(step3.n, starts_with("sample")), 2, sum)) # 997
mean(apply(select(step3.n, starts_with("sample")), 2, sum)) # 166.1667
sd(apply(select(step3.n, starts_with("sample")), 2, sum)) # 337.9073
length(apply(select(step3.n, starts_with("sample")), 2, sum)) # 6
337.9073/sqrt(6) # 137.9501


### Coleop

# Coleop step 1: how many reads after filtering steps performed on cluster

step1 <- c %>%
  left_join(select(c2, id, definition, sequence, starts_with("sample")), by = join_by(id, definition, sequence))

# check which columns are negatives
colnames(step1) # input those column numbers in the next line
negatives.C <- c(15,16,24,25,56:68)
colnames(as.data.frame(step1)[, negatives.C])

# select out just the samples
step1.s <- step1 %>%
  select(-all_of(negatives.C))

# tally up read stats for samples
sum(apply(select(step1.s, starts_with("sample")), 2, sum, na.rm = T)) # 2668687
mean(apply(select(step1.s, starts_with("sample")), 2, sum, na.rm = T)) # 72126.68
sd(apply(select(step1.s, starts_with("sample")), 2, sum, na.rm = T)) # 31761.21
length(apply(select(step1.s, starts_with("sample")), 2, sum)) # 37
31761.21 / sqrt(37) # 5221.511

# select out just the negatives
step1.n <- step1 %>%
  select(id, all_of(negatives.C), sequence)

# tally up read stats for negatives
sum(apply(select(step1.n, starts_with("sample")), 2, sum)) # 60072
mean(apply(select(step1.n, starts_with("sample")), 2, sum)) # 3533.647
sd(apply(select(step1.n, starts_with("sample")), 2, sum)) # 8229.316
length(apply(select(step1.n, starts_with("sample")), 2, sum)) # 17
8229.316 / sqrt(17) # 1995.902

# Coleop step 2: remove human and mouse 

step2.s <- step1.s %>%
  left_join(select(cs, id:species)) %>%
  select(id, match, finalID:species, starts_with("sample")) %>%
  filter(species %notin% c("Homo sapiens", "Mus musculus")) # no mice or humans in COI primer

# tally up read stats for samples
sum(apply(select(step2.s, starts_with("sample")), 2, sum)) # 2611101
mean(apply(select(step2.s, starts_with("sample")), 2, sum, na.rm = T)) # 70570.3
sd(apply(select(step2.s, starts_with("sample")), 2, sum, na.rm = T)) # 30645.44
length(apply(select(step2.s, starts_with("sample")), 2, sum)) # 37
30645.44 / sqrt(37) # 5038.079

step2.n <- step1.n %>%
  left_join(select(cs, id:species)) %>%
  select(id, finalID:species, starts_with("sample")) %>%
  filter(species %notin% c("Homo sapiens", "Mus musculus")) 

# tally up read stats for negatives
sum(apply(select(step2.n, starts_with("sample")), 2, sum)) # 6413
mean(apply(select(step2.n, starts_with("sample")), 2, sum, na.rm = T)) # 377.2353
sd(apply(select(step2.n, starts_with("sample")), 2, sum, na.rm = T)) # 986.0064
length(apply(select(step2.n, starts_with("sample")), 2, sum)) # 17
986.0064 / sqrt(17) # 239.1417

# Coleop step 3: subtract negatives, remove rest of non-arthropods
  # this is same as zs3 from above
step3.s <- cs3

# tally up read stats for samples
sum(apply(select(step3.s, starts_with("sample")), 2, sum)) # 2240233
mean(apply(select(step3.s, starts_with("sample")), 2, sum)) # 60546.84
sd(apply(select(step3.s, starts_with("sample")), 2, sum)) # 32800.04
length(apply(select(step3.s, starts_with("sample")), 2, sum)) # 37
32800.04 / sqrt(37) # 5392.293

step3.n <- cn3

# tally up read stats for samples
sum(apply(select(step3.n, starts_with("sample")), 2, sum)) # 4442
mean(apply(select(step3.n, starts_with("sample")), 2, sum)) # 261.2941
sd(apply(select(step3.n, starts_with("sample")), 2, sum)) # 802.3041
length(apply(select(step3.n, starts_with("sample")), 2, sum)) # 17
802.3041/sqrt(17) # 194.5873
```
# Negative control read tally
How many negative control samples with reads, etc.
```{r}

# list all negative control samples
unique(c(names(cn3)[12:28], names(zn3)[12:17]))

# how many negative MOTUs were also in samples
# COI (Z): 
test <- zn3[zn3$seqmax>0,] # get all the arthropod MOTUs with reads in the negatives

# see the MOTUs in the samples that had reads in the negatives after filtering steps
test2 <- zs3 %>%
  filter(id %in% zn3[zn3$seqmax>0,]$id,
         sampsum > 0)

# calculate stats for "contaminant" MOTUs that also appear in samples
test3 <- zn3 %>%
  filter(id %in% test2$id)
sum(test3$negsum); min(test3$negsum); max(test3$negsum)

# 16S (C): 
test <- cn3[cn3$seqmax>0,] # get all the MOTUs with reads in the negatives

# see the MOTUs in the samples that had reads in the negatives after filtering steps
test2 <- cs3 %>%
  filter(id %in% test$id,
         sampsum > 0)

# calculate stats for "contaminant" MOTUs that also appear in samples
test3 <- cn3 %>%
  filter(id %in% test2$id)
sum(test3$negsum); min(test3$negsum); max(test3$negsum)

## count other taxa in initial MOTUs
table(cc_finID$phylum)
table(zz_finID$phylum)
table(cc_finID$class)
table(zz_finID$class)

# look at MOTUs with no phylum
testC <- cc_finID[is.na(cc_finID$phylum),]
testZ <- zz_finID[is.na(zz_finID$phylum),]
```

